---
title: "High-Protein BigCages - Fecundity"
Author: "Claudia Ramirez-Lanzas"
Date: "06/02/2025"
---

# Load libraries:
```{r}
# Data handling:
library("readxl")
library("writexl")
library("xlsx")
library("tidyverse")

# Statistics:
library("lme4")
library("lmerTest") # to include approximate p-values in the output of lmer. 
library("emmeans")
library("car")
library("HLMdiag")
library("ggeffects")

# Plotting:
library("patchwork")
library("ggplot2")
library("ggpubr") 


```

```{r}
sessionInfo()
```
# Import files:
```{r}
fecundity_Big <- read_csv("input_files/BigCages_fecundity_weight_per_transfer.csv", show_col_types = FALSE)
```

```{r}
str(fecundity)
```
# ####Part 1: investigate raw data
# Designate variables as a categorical and numerical factors:
```{r}
# Format the variables:
cols_factor <- c("CGE","generation","day","transfer", "rep", "subrep", 
                 "subsubrep", "treatment")
cols_numeric <- c("nb_eggs", "nb_females", "nb_males", "total_flies",
                  "females_mg", "females_g", "males_mg", "males_g")
fecundity_Big[cols_factor] <- lapply(fecundity_Big[cols_factor], as.factor)
fecundity_Big[cols_numeric] <- lapply(fecundity_Big[cols_numeric], as.numeric)

# re-order factor levels for treatment:
fecundity_Big$treatment <- factor(fecundity_Big$treatment, 
                                   levels=c("base", "trial_big"))

# re-order factor levels for generation:
fecundity_Big$generation <- factor(fecundity_Big$generation, 
                                    levels=c("gen0", "gen7", "gen31"))

# Include subsubrep for trial_big replicates:
fecundity_Big <- fecundity_Big %>% dplyr::mutate(subsubrep = replace_na(subsubrep, "1"))

```
# Normalize fecundity by number of females and dry weight:
```{r}
fecundity_Big_transfer <- fecundity_Big 
fecundity_Big_transfer$eggs_female <- fecundity_Big_transfer$nb_eggs / fecundity_Big_transfer$nb_females
fecundity_Big_transfer$eggs_mg <- fecundity_Big_transfer$nb_eggs / fecundity_Big_transfer$females_mg

```
# Compute total fecundity (7 days) and normalize by nb_females and total_mg:
```{r}
# Aggregate data so number of females is added for 7 days.
fecundity_Big_total <- aggregate(nb_eggs ~ CGE + 
                                   generation + 
                                   treatment + 
                                    rep + 
                                   subrep + 
                                   subsubrep + 
                                   nb_females + 
                                   nb_males +
                                   total_flies +
                                   females_mg,
                                  data = as.data.frame(fecundity_Big), 
                                  FUN = sum, na.rm = TRUE)


# Compute mg dry weight per female:
fecundity_Big_total$mg_female <- fecundity_Big_total$females_mg / fecundity_Big_total$nb_females
# Compute fecundity_total by dividing total nb_eggs in 7 days per nb_females:
fecundity_Big_total$eggs_female <- fecundity_Big_total$nb_eggs / fecundity_Big_total$nb_females
# Compute fecundity_total by dividing total nb_eggs in 7 days per total dry weight:
fecundity_Big_total$eggs_mg <- fecundity_Big_total$nb_eggs / fecundity_Big_total$females_mg

```
# Export BigCages_fecundity_total
```{r}
export_df <- fecundity_Big_total

#write_csv(export_df, "input_files/BigCages_fecundity_total.csv")
  
```


# Compute daily fecundity:
```{r}
# Aggregate data so number of eggs is added for each day:
fecundity_day_Big <- aggregate(eggs_female ~ CGE + 
                                   generation + 
                                   treatment + 
                                    rep + 
                                   subrep + 
                                   subsubrep + 
                                   nb_females + 
                                   females_mg + day, 
                                  data = as.data.frame(fecundity_Big_transfer), 
                                  FUN = sum, na.rm = TRUE)
```
# Compute total fecundity (7 days) and normalize by nb_females and total_mg:
```{r}
# Aggregate data so number of females is added for 7 days.
fecundity_Big_total <- aggregate(nb_eggs ~ CGE + 
                                   generation + 
                                   treatment + 
                                    rep + 
                                   subrep + 
                                   subsubrep + 
                                   nb_females + 
                                   females_mg,
                                  data = as.data.frame(fecundity_Big), 
                                  FUN = sum, na.rm = TRUE)

# Compute mg dry weight per female:
fecundity_Big_total$mg_female <- fecundity_Big_total$females_mg / fecundity_Big_total$nb_females
# Compute fecundity_total by dividing total nb_eggs in 7 days per nb_females:
fecundity_Big_total$eggs_female <- fecundity_Big_total$nb_eggs / fecundity_Big_total$nb_females
# Compute fecundity_total by dividing total nb_eggs in 7 days per total dry weight:
fecundity_Big_total$eggs_mg <- fecundity_Big_total$nb_eggs / fecundity_Big_total$females_mg

```
## Remove 2 outliers in Big Cages - total and daily datasets
We have evidences that CGE2_2_35_2 & CGE2_2_35_3 samples should be removed: our
notebook indicates that many flies escaped at once during the fecundity assay
meaning that the fecundity measurements of previous days will be overestimated.
```{r}
# Create unique ID:
fecundity_Big_total$unique_id <- paste(fecundity_Big_total$CGE, fecundity_Big_total$rep, fecundity_Big_total$subrep, sep="_")
# Remove samples:
fecundity_Big_filter <- fecundity_Big_total %>% dplyr::filter(!unique_id %in% c("CGE2_2_35_2", "CGE2_2_35_3"))

# Create unique ID:
fecundity_day_Big$unique_id <- paste(fecundity_day_Big$CGE, fecundity_day_Big$rep, fecundity_day_Big$subrep,  sep="_")
# Remove samples:
fecundity_day_Big_filter <- fecundity_day_Big %>% dplyr::filter(!unique_id %in% c("CGE2_2_35_2", "CGE2_2_35_3"))

```
# Plot - total fecundity
```{r}
# Fecundity normalized by number of females:
pallete_treatments <- c("orange","#af8dc3","#762a83")


fecundity_Big_filter %>%
  ggplot(aes(x=generation, y=eggs_female, fill=generation))  +
  geom_boxplot(outlier.size = 0.0001) + 
  geom_text(data = fecundity_Big_filter %>% filter(!is.na(eggs_female)), 
            aes(label = ifelse(eggs_female > quantile(eggs_female, 0.75) + 1.5 * IQR(eggs_female) | 
                               eggs_female < quantile(eggs_female, 0.25) - 1.5 * IQR(eggs_female), 
                             as.character(rep), "")),
            position = position_dodge(width = 0.75), vjust = -0.5) + 
  labs(title="Total fecundity per generation") + 
  scale_fill_manual(values = pallete_treatments) +  
  ylab("Fecundity (No eggs / fly / 7 days)") + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  guides(fill = FALSE) +
  theme_light() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 13)) + 
  xlab("")

```
# Plot - daily fecundity
```{r}
# Fecundity normalized by number of females:
pallete_treatments <- c("orange","#af8dc3","#762a83")


plot_daily_fec <- fecundity_day_Big_filter %>%
  ggplot(aes(x = day, y = eggs_female, fill = generation))  +
  geom_boxplot(outlier.size = 0.0001, show.legend = TRUE) +
  scale_fill_manual(values = pallete_treatments,
                    labels = c("Gen 0", "Gen 7", "Gen 31"),
                    name ="") +  
  ylab("Fecundity (No eggs / fly)") + 
  theme_classic() +
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        text = element_text(size = 14)) + 
  xlab("") +
  scale_x_discrete(labels = c("1","2","3","4","5","6","7")) + 
  xlab("Day")

plot_daily_fec

# Export plot:
#ggsave("output_files/plots/High_order_phenotypes/final_version/fecundity_daily_raw_data.png",
       #plot_daily_fec,
       #dpi=400,
       #units = "mm",
        #height = 150, width=150)

# Print mean:
fecundity_day_Big_filter %>% group_by(generation, day) %>%
  summarize(mean_fecundity = mean(eggs_female),
            median_fecundity= median(eggs_female), max_fec = max(eggs_female))

```
# ####Part 2: overall model  
Here we test the difference in fecundity across generations 0, 7 and 31.
## Create model_dataset:
```{r}
model_dataset <- fecundity_Big_filter

# Homogenize labeling:
model_dataset$subrep <- sub("^0", "", model_dataset$subrep)
model_dataset$subsubrep <- sub("^0", "", model_dataset$subsubrep)
model_dataset <- model_dataset %>%
  mutate(subsubrep = case_when(subsubrep == "NA" ~ "1", TRUE ~ subsubrep))


#Generations random slopes:
### Dummy coding and centering of random slope categorical variables: 
model_dataset$gen_7_dummy <- as.numeric(
  model_dataset$generation == levels(model_dataset$generation)[2]) # Dummy codding for 2nd level
model_dataset$gen_7_dummy <- model_dataset$gen_7_dummy - 
  mean(model_dataset$gen_7_dummy) # Centering
model_dataset$gen_31_dummy <- as.numeric(
  model_dataset$generation == levels(model_dataset$generation)[3]) # Dummy codding for 3rd level
model_dataset$gen_31_dummy <- model_dataset$gen_31_dummy - 
  mean(model_dataset$gen_31_dummy) # Centering

# Format the variables:
cols_factor <- c("CGE",
                 "rep", 
                 "subrep", 
                 "generation",
                 "unique_id")
model_dataset[cols_factor] <- lapply(model_dataset[cols_factor], as.factor)

model_dataset$generation <- factor(model_dataset$generation,
                            levels = c("gen0", "gen7","gen31"))

head(model_dataset)

```

## Fit the model:
```{r}

model_1 <-  lmer(log10(eggs_female) ~ generation + mg_female + (1 + gen_7_dummy + gen_31_dummy || rep), data = model_dataset, REML = FALSE)
summary(model_1)

# Introduce the residuals in the data-set:
model_dataset_1 <- model_dataset
model_dataset_1$residuals <- residuals(model_1)

```
1 + Allows for a random intercept per rep, meaning each replicate can have
a different baseline level of eggs_female.
generation_7_dummy and generation_31_dummy: allow the slopes of gen 7 and gen 31
to vary by rep, meaning the effect of  generations can differ across replicates.

## Check model assumptions:
```{r}
# Homogeneity of variance:
#Predicted values are plotted against residuals to assess variance homogeneity.
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ generation, data = model_dataset_1)

# Normality of residuals:
# Q-Q plot of normal residuals:
qqPlot(residuals(model_1))
# Normality test 1:
#Shapiro-Wilk’s method is widely recommended for normality test 
#and it provides better power than K-S. It is based on the 
#correlation between the data and the corresponding normal scores.
shapiro.test(residuals(model_1)) 

# Check assumptions for random intercept term:
# Are random intercept normally distributed?
ranef_rep <- ranef(model_1)[[1]]
ranef_rep$rep <- rownames(ranef_rep)
ranef_rep <- merge(ranef_rep, model_dataset_1[!duplicated(model_dataset_1$rep), 
                                              c("generation","CGE", "rep")], by = "rep")

qqnorm(ranef_rep$`(Intercept)`, 
       main="QQ Plot for random intercepts per Generation", 
       col = ranef_rep$generation, pch =19)

```
# Contrast - generations
```{r}

contrasts <- summary(emmeans(model_1, pairwise ~ generation, 
                      type="response", adjust="tukey")$contrasts)

# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ generation, 
                      type="response")$emmeans) 

# Make generation a continuous variable for plotting:
predicted_means <- predicted_means %>% dplyr::mutate(generation = case_when(generation == "gen0" ~ 0, 
                                                                            generation == "gen7" ~ 7,
                                                                            generation == "gen31" ~ 31))

# Change generation scale to reduce distances proportionally for plotting:
predicted_means <- predicted_means %>% dplyr::mutate(generation_label = case_when(generation == "0" ~ 0.0, 
                                                                            generation == "7" ~ 0.5,
                                                                            generation == "31" ~ 1.5))

# Visualize the emmeans:
plot(emmeans(model_1, pairwise ~ generation, 
                      type="response", adjust="tukey"))

## Make a data-set with results:
results <- as.data.frame(contrasts)
results <- results %>%
  separate(contrast,
           into = c("group1", "group2"), sep = " / ")

results

# Export results:
#write.xlsx(results, file = "output_files/S4_models_high_order_phenotypes.xlsx", sheetName="fec_overall", append=TRUE)

```
# Plot - emmeans
```{r}

plot_overall_fec <- predicted_means %>% ggplot(
  aes(x = (generation_label),
      y = response, fill = as.factor(generation_label))) + 
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.4),
           width=0.3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.3),  
                width = 0.1) + 
  scale_fill_manual(values = c("#f1a340","#af8dc3", "#762a83")) +  
  labs(title = "",
       x = "",
       y = "Fecundity (No. eggs / fly / 7 days)") +  
  theme_classic() +
  xlab("Generation") + 
  theme(aspect.ratio = 1, plot.margin = margin(0, 0, 0, 0)) + 
  theme(axis.text.x.top  = element_blank(),
        #axis.line.x.top = element_blank(),
        #axis.ticks.x.top = element_blank(),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="none") +
  coord_cartesian(ylim = c(0, 100)) +
  scale_x_continuous(breaks = c(0, 0.5, 1.5),
                     labels = c(0,7,31))


plot_overall_fec 

# Export plot:
#ggsave("output_files/plots/High_order_phenotypes/fecundity_overall_emmeans.svg",
#plot_overall_fec, units = "mm", height = 150, width = 150)


```

# ####Part 3: model replicate-wise model 
Here we test the difference in fecundity across the 6 different replicates
at generation 7 and at generation 31 by fitting one model for each generation.
## Create  datasets for F7 and F31:
```{r}
model_dataset <- fecundity_Big_filter

# Homogenize labeling:
model_dataset$subrep <- sub("^0", "", model_dataset$subrep)
model_dataset$subsubrep <- sub("^0", "", model_dataset$subsubrep)
model_dataset <- model_dataset %>% mutate(subsubrep = case_when(subsubrep == "NA" ~ "1", TRUE ~ subsubrep))

# Subset data-sets for the F7 models:
big_F7 <- model_dataset %>% filter(generation == "gen7") 
# Format the variables:
cols_factor <- c("rep", 
                 "subrep", 
                 "subsubrep", "generation", "CGE")

big_F7[cols_factor] <- lapply(big_F7[cols_factor], as.factor)

## 
# Subset data-sets for the F31 models:
big_F31 <- model_dataset %>% filter(generation == "gen31") 
# Format the variables:
cols_factor <- c("rep", 
                 "subrep", 
                 "subsubrep", "generation", "CGE")

big_F31[cols_factor] <- lapply(big_F31[cols_factor], as.factor)

```
# Fit the model - F7:
We fit a simple linear model with replicate as a fixed categorical variable for
replicate and fixed continuous variable for weight, since we have added weight
in the overall model. 
```{r}

# Fit the model:
model_1 <- lm(log10(eggs_female) ~ rep + mg_female, data = big_F7) 
summary(model_1)

#Create model_dataset_1
model_dataset_1 <- big_F7
## Introduce the residuals in the dataset:
model_dataset_1$residuals <- residuals(model_1)

```
# Check model assumptions:
```{r}

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_dataset_1)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 


```
# Remove 2 outliers and fit the model:
```{r}
model_dataset_1 <- model_dataset_1 %>% dplyr::filter(unique_id != c("CGE2_2_34_3", "CGE2_2_33_2"))

model_1 <- lm(log10(eggs_female) ~ rep + mg_female, data = model_dataset_1) 
summary(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_dataset_1)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Contrasts:
```{r}
contrasts <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="none")$contrasts)

# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="tukey")$emmeans) 

predicted_means

## Make a data-set with results:
results_F7 <- as.data.frame(contrasts)
results_F7 <- results_F7 %>%
  separate(contrast, into = c("group1", "group2"), sep = " / ")

results_F7


# Visualize the emmeans:
plot(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="tukey"))

```
# Plot - emmeans F7
```{r}
pallete_treatments <- c("#af8dc3","#af8dc3",
                        "#af8dc3","#af8dc3",
                        "#af8dc3","#af8dc3")

plot_F7 <- ggplot(predicted_means, aes(x = rep, y = response, fill = rep)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.9),
                width = 0.25) + 
  scale_fill_manual(values = pallete_treatments) +  
  labs(title = "Generation 7",
       x = "Replicate",
       y = "Fecundity (No eggs / fly / 7 days)") + 
  scale_x_discrete(labels = (c("1", "2", "3", "4", "5", "6"))) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 20)) + 
  theme(aspect.ratio = 1) +  
  theme(plot.margin = margin(0, 0, 0, 0)) + theme(
    axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color="black"), 
    axis.line = element_line(color = "black")) + coord_cartesian(ylim = c(0,100))


plot_F7

```
# Fit the model - F31:
```{r}

# Fit simpler model:
model_1 <- lm(log10(eggs_female) ~ rep + mg_female, data = big_F31) 
summary(model_1)
#Create model_dataset_1
model_dataset_1 <- big_F31
## Introduce the residuals in the dataset:
model_dataset_1$residuals <- residuals(model_1)

```
# Check model assumptions:
```{r}

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_dataset_1)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Remove 1 outlier and fit the model:
```{r}
model_dataset_1 <- model_dataset_1 %>% dplyr::filter(unique_id != "CGE3_2_34_1")

model_1 <- lm(log10(eggs_female) ~ rep + mg_female, data = model_dataset_1) 
summary(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_dataset_1)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Contrasts:
```{r}
contrasts <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="none")$contrasts) 
# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="tukey")$emmeans) 

predicted_means

## Make a data-set with results:
results_F31 <- as.data.frame(contrasts)
results_F31 <- results_F31 %>%
  separate(contrast, into = c("group1", "group2"), sep = " / ")

results_F31


# Visualize the emmeans:
plot(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="tukey"))

```
# Plot - emmeans F7
```{r}
pallete_treatments <- c("#762a83","#762a83",
                        "#762a83","#762a83",
                        "#762a83","#762a83")

plot_F31 <- ggplot(predicted_means, aes(x = rep, y = response, fill = rep)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.9),
                width = 0.25) + 
  scale_fill_manual(values = pallete_treatments) +  
  labs(title = "Generation 31",
       x = "Replicate",
       y = "") + #Fecundity (No eggs / fly / 7 days)
  scale_x_discrete(labels = (c("1", "2", "3", "4", "5", "6"))) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 20)) + 
  theme(aspect.ratio = 1) +  
  theme(plot.margin = margin(0, 0, 0, 0)) + theme(
    axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color="black"), 
    axis.line = element_line(color = "black")) + coord_cartesian(ylim = c(0,100))


plot_F31
```
# Plot emmeans - F7 F31
```{r}
# Combine the plots using patchwork
combined_plot <- plot_F7 + plot_F31 & 
  theme(axis.title.x = element_blank(),  
    axis.text.x = element_text(size = 16),  
    axis.ticks.x = element_line()) 

combined_plot_rep_fec <- wrap_elements(panel = combined_plot) +
  labs(tag = "Replicate") +
  theme(
    plot.tag = element_text(size = 16),
    plot.tag.position = "bottom")

combined_plot_rep_fec


# Export plot:
#ggsave("output_files/plots/High_order_phenotypes/fecundity_repwise_emmeans.svg",
#combined_plot_rep_fec, units = "mm", height = 150, width = 150)

```
# Final p-values: correction across all contrasts
The p-value of emmeans was not corrected (p.value), here we correct across all
contrasts using FDR.
```{r}
results_F7$group <- "F7"
results_F31$group <- "F31"

combined_contrasts_fecundity_rep <- rbind(results_F7,results_F31)

# Correct for multiple testing across all contrasts and sex:
combined_contrasts_fecundity_rep$padj_fdr <- p.adjust(combined_contrasts_fecundity_rep$p.value, method = "fdr", n = length(combined_contrasts_fecundity_rep$p.value))
combined_contrasts_fecundity_rep

# Export results:
#write.xlsx(results, file = "output_files/S4_models_high_order_phenotypes.xlsx", sheetName="fec_repwise", append=TRUE)

```

--- END ---