---
title: "High-Protein BigCages - Dry-weight"
Author: "Claudia Ramirez-Lanzas"
Date: "11/06/2024"
---

# Load libraries:
```{r}
# Data handling:
library("readxl")
library("writexl")
library("xlsx")
library("tidyverse")

# Statistics:
library("lme4")
library("lmerTest") # to include approximate p-values in the output of lmer. 
library("emmeans")
library("car")
library("HLMdiag")
library("ggeffects")

# Plotting:
library("patchwork")
library("ggplot2")
library("ggpubr") 

```

```{r}
sessionInfo()
```
# Part 1: Prepare the data-set and plot raw data
# Import data-set:
```{r}
dry_weight_Big <- read_csv("input_files/BigCages_dry_weight.csv", show_col_types = FALSE) 

```
# Designate variables as a categorical and numerical factors:
```{r}
# Format the variables:
cols_factor <- c("CGE","generation", "rep", "subrep", 
                 "subsubrep", "treatment")

cols_numeric <- c("nb_females", "nb_males", "total_flies",
                  "females_mg", "females_g", "males_mg", "males_g")
dry_weight_Big[cols_factor] <- lapply(dry_weight_Big[cols_factor], as.factor)
dry_weight_Big[cols_numeric] <- lapply(dry_weight_Big[cols_numeric], as.numeric)

# re-order factor levels for treatment:
dry_weight_Big$treatment <- factor(dry_weight_Big$treatment, 
                                   levels=c("base", "trial_big"))
# re-order factor levels for generation:
dry_weight_Big$generation <- factor(dry_weight_Big$generation, 
                                    levels=c("gen0", "gen7", "gen31"))

```
# Check data-set structure:
```{r}
str(dry_weight_Big)
```
# Prepare variables:
Dry weight was measured as the mass of all females and males
in each replicate used for the fecundity assay. 
```{r}
# Keep needed columns:
dry_weight_Big <- dry_weight_Big %>% dplyr::select(c("CGE","rep","subrep",
                                                   "subsubrep","generation",
                                                   "treatment",
                                                   "nb_females","nb_males",
                                                   "total_flies", "females_mg",
                                                   "males_mg"))
# Homogenize labeling:
dry_weight_Big$subrep <- sub("^0", "", dry_weight_Big$subrep)
dry_weight_Big$subsubrep <- sub("^0", "", dry_weight_Big$subsubrep)


table(dry_weight_Big$treatment, dry_weight_Big$CGE)
table(dry_weight_Big$treatment)
table(dry_weight_Big$generation)
table(dry_weight_Big$CGE)

```
# Compute mg per fly:
```{r}

dry_weight_Big$female <- dry_weight_Big$females_mg / dry_weight_Big$nb_females
dry_weight_Big$male <- dry_weight_Big$males_mg / dry_weight_Big$nb_males

# Long format:
dry_weight_Big_long <-  dry_weight_Big %>%
  tidyr::pivot_longer(cols = c(female, male),
                      names_to = "sex",
                      values_to = "mg_fly")

```

# Plot -  dry weight per sex
```{r}
# Fecundity normalized by number of females:
pallete_treatments <- c("orange","#af8dc3","#762a83")

dry_weight_Big_long %>%
  ggplot(aes(x = generation,
             y = mg_fly,
             fill = generation))  +
  geom_boxplot(outlier.size = 0.0001) + 
  geom_text(data = dry_weight_Big_long %>% filter(!is.na(mg_fly)), 
            aes(label = ifelse(mg_fly > quantile(mg_fly, 0.75) + 1.5 * IQR(mg_fly) | 
                               mg_fly < quantile(mg_fly, 0.25) - 1.5 * IQR(mg_fly), 
                             as.character(rep), "")),
            position = position_dodge(width = 0.75), vjust = -0.5) + 
  scale_fill_manual(values = pallete_treatments) +  
  ylab("mg dry weight / fly") + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  guides(fill = FALSE) +
  theme_light() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 13)) + 
  xlab("") +
  facet_wrap(~sex)

plot_df <- dry_weight_Big_long

plot_df %>% ggplot(aes(x = as.numeric(generation), y = mg_fly, color = rep)) +
  geom_point(size=4, alpha=0.4) + facet_wrap(~ sex)
  
```
# Plot - dry weight per replicate and sex
```{r}
# Fecundity normalized by number of females:
pallete_treatments <- c("orange","#af8dc3","#762a83")

dry_weight_Big_long %>%
  ggplot(aes(x = rep,
             y= mg_fly,
             fill=generation))  +
geom_boxplot() +
  scale_fill_manual(values = pallete_treatments) +  
  ylab("mg dry weight / fly") + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  guides(fill = FALSE) +
  theme_light() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 13)) + 
  xlab("") +
  facet_wrap(~sex+generation)



```
## Remove 2 outliers:
We have evidences that CGE2_2_35_2 & CGE2_2_35_3 samples should be removed: our
notebook indicates that many flies escaped at once during the fecundity assay so
those samples are removed for this analysis and also the fecundity analysis.
```{r}
# Create unique ID:
dry_weight_Big_long$unique_id <- paste(dry_weight_Big_long$CGE, dry_weight_Big_long$rep, dry_weight_Big_long$subrep, sep="_")

# Remove sample CGE2_2_35_3
dry_weight_Big_filter <- dry_weight_Big_long %>% dplyr::filter(!unique_id %in% c("CGE2_2_35_3", "CGE2_2_35_2"))

```



# ####Part 2: - overall model 
Here we test the difference in body weight across generations 0, 7 and 31 for 
females and males.
# #Create model dataset:

Here we aggregate the replicate weights using the average of the 3 sub-replicates
because we were not able to make the residuals homogeneous in previous analysis.
```{r}
model_dataset_full <- dry_weight_Big_filter %>% dplyr::select(c( "CGE", "generation",
                                                        "rep", "subrep",
                                                        "treatment",
                                                        "sex", 
                                                 "mg_fly"))
  
  
# The samples from base population are not averaged:
base <- model_dataset_full %>% dplyr::filter(CGE == "CGE1")

# The samples from evolved populations are averaged within each replicate:
evolved <- model_dataset_full %>% dplyr::filter(CGE != "CGE1")
evolved_aggre <- evolved %>% 
  group_by(generation, CGE,rep, sex) %>% 
  summarise(mg_fly = mean(mg_fly))

# Select columns to match with evolved_aggre:
base <- base %>% dplyr::select(c(colnames(evolved_aggre)))

# Combine base and aggregated evolved replicates:
model_dataset_aggre <- rbind(base, evolved_aggre)

# Format the variables:
cols_factor <- c("rep", 
                 "generation",
                 "sex")

model_dataset_aggre[cols_factor] <- lapply(model_dataset_aggre[cols_factor], as.factor)

table(model_dataset_aggre$rep, model_dataset_aggre$generation)


```

# Fit the model
```{r}
model_1 <-  lmer((mg_fly) ~ generation*sex + (1 | rep),
                 data = model_dataset_aggre,
                 REML = FALSE)
summary(model_1)

# Introduce the residuals in the data-set:
model_dataset_1 <- model_dataset_aggre
model_dataset_1$residuals <- residuals(model_1)
model_dataset_1$unique_id <- paste(model_dataset_1$CGE, model_dataset_1$rep, model_dataset_1$sex, sep="_")

```
# Check model assumptions:
```{r}

# Homogeneity of variance:
#Predicted values are plotted against residuals to assess variance homogeneity.
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. # Homogeneous variance P-value= 
leveneTest(residuals ~ generation, data = model_dataset_1)


# Normality of residuals:
# Q-Q plot of normal residuals:
qqPlot(residuals(model_1))

# Normality test 1:
shapiro.test(residuals(model_1))
# Normality test 2:
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

# Check assumptions for random intercept term:
# Are random intercept normally distributed?
ranef_rep <- ranef(model_1)[[1]]
ranef_rep$rep <- rownames(ranef_rep)
ranef_rep <- merge(ranef_rep, model_dataset_1[!duplicated(model_dataset_1$rep), 
                                              c("generation", "rep")], by = "rep")

qqnorm(ranef_rep$`(Intercept)`, 
       main="QQ Plot for random intercepts per generation", 
       col = ranef_rep$generation, pch =19)


```
# Remove 2 outliers and fit the model:
```{r}
model_dataset_1 <- model_dataset_1 %>% dplyr::filter(residuals < 0.04) %>% dplyr::filter(residuals > -0.05)

model_1 <-  lmer((mg_fly) ~ generation*sex + (1 | rep), data = model_dataset_1, REML = FALSE)
summary(model_1)

# Introduce the residuals in the data-set:
model_dataset_1$residuals <- residuals(model_1)
model_dataset_1$unique_id <- paste(model_dataset_1$CGE,
                                   model_dataset_1$rep,
                                   model_dataset_1$sex,
                                   sep="_")

```
# Check model assumptions:
```{r}

# Homogeneity of variance:
#Predicted values are plotted against residuals to assess variance homogeneity.
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. # Homogeneous variance P-value= 
leveneTest(residuals ~ generation, data = model_dataset_1)

# Normality of residuals:
# Q-Q plot of normal residuals:
qqPlot(residuals(model_1))
# Normality test 1:
shapiro.test(residuals(model_1))
# Normality test 2:
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

# Check assumptions for random intercept term:
# Are random intercept normally distributed?
ranef_rep <- ranef(model_1)[[1]]
ranef_rep$rep <- rownames(ranef_rep)
ranef_rep <- merge(ranef_rep, model_dataset_1[!duplicated(model_dataset_1$rep), 
                                              c( "generation", "rep")], by = "rep")

qqnorm(ranef_rep$`(Intercept)`, 
       main="QQ Plot for random intercepts per generation", 
       col = ranef_rep$generation, pch =19)


```
# Contrasts:
```{r}
# Contrasts:
contrasts <- summary(emmeans(model_1,
                             pairwise ~ generation|sex, 
                      type="response",
                      adjust = "none")$contrasts) 
contrasts

# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ generation|sex, 
                      type="response", adjust="none")$emmeans) 

# Change generation scale to reduce distances proportionally: (make nicer plot)
predicted_means <- predicted_means %>% dplyr::mutate(generation = case_when(generation == "gen0" ~ 0, 
                                                                            generation == "gen7" ~ 0.5,
                                                                            generation == "gen31" ~ 1.5))

# Plot:
plot(emmeans(model_1, pairwise ~ generation|sex, 
                      type="response", adjust = "none"))

## Make a data-set with results:
results <- as.data.frame(contrasts)
results <- results %>%
  separate(contrast, into = c("group1", "group2"), sep = " - ")

results

```
# Plot - emmeans
```{r}

pallete_treatments <- c("#f1a340","#af8dc3", "#762a83")

plot1 <- predicted_means %>% ggplot(
  aes(x = generation,
      y = response, fill=as.factor(generation))) + 
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.4),
           width=0.3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.3),
                width = 0.1) + 
  scale_fill_manual(values = pallete_treatments) + 
  facet_wrap(~sex, 
             labeller = as_labeller(
               c(female = "Females", male = "Males"))) +
  labs(title = "",
       x = "",
       y = "Dry weight (mg / fly)") +  
  theme_classic() +
  xlab("Generation") + 
  theme(aspect.ratio = 1, plot.margin = margin(0, 0, 0, 0)) + 
  theme(axis.text.x.top  = element_blank(),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size = 20, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    strip.background = element_blank(),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 16, face = "bold"),
    legend.position="none") +
  scale_x_continuous(breaks = c(0, 0.5, 1.5),
                     labels = c(0,7,31))

plot1

# Export plot:
#ggsave("output_files/plots/High_order_phenotypes/body_weight_overall_emmeans.svg",
#plot_overall_fec, units = "mm", height = 160, width = 300)

```
# Final p-values: correction across all contrasts
The p-value of emmeans was not corrected (p.value)
```{r}

# Correct for multiple testing across all contrasts and sex:
results$padj_fdr <- p.adjust(results$p.value, method = "fdr",
                             n = length(results$p.value))
results


# Export results:
#write.xlsx(results, file = "output_files/S4_models_high_order_phenotypes.xlsx", sheetName="weight_overall", append=TRUE)

```

# #### Part 3: replicate-wise model
Here we test the difference in body weight across the 6 different replicates
at generation 7 and at generation 31 by fitting one model for each generation
and sex.
## Create  datasets for F7 and F31:
```{r}

# Subset data-sets for the 4 models:
big_F7_females <- dry_weight_Big_filter %>% filter(generation == "gen7" & sex == "female") 
# Format the variables:
cols_factor <- c("rep", 
                 "subrep", 
                 "subsubrep")
big_F7_females[cols_factor] <- lapply(big_F7_females[cols_factor], as.factor)


##
big_F7_males <- dry_weight_Big_filter %>% filter(generation == "gen7" & sex == "male") # 
# Format the variables:
cols_factor <- c("rep", 
                 "subrep", 
                 "subsubrep")
big_F7_males[cols_factor] <- lapply(big_F7_males[cols_factor], as.factor)


##
big_F31_females <- dry_weight_Big_filter %>% filter(generation == "gen31" & sex == "female") 
# Format the variables:
cols_factor <- c("rep", 
                 "subrep", 
                 "subsubrep")
big_F31_females[cols_factor] <- lapply(big_F31_females[cols_factor], as.factor)

##
big_F31_males <- dry_weight_Big_filter %>% filter(generation == "gen31" & sex == "male") 
# Format the variables:
cols_factor <- c("rep", 
                 "subrep", 
                 "subsubrep")
big_F31_males[cols_factor] <- lapply(big_F31_males[cols_factor], as.factor)

```
# #F7  Females
# Fit the model 
```{r}
#model_1 <- lmer(log10(mg_fly) ~ rep + (1 | rep), data = big_F7_females) # Warning: Model may not have converged with 1 eigenvalue close to zero: 9.2e-10. 

model_1 <- lm(log10(mg_fly) ~ rep, data = big_F7_females) 

summary(model_1)

```
# Check model assumptions:
```{r}
#Create model_1_dataset
model_1_dataset <- big_F7_females
## Introduce the residuals in the dataset:
model_1_dataset$residuals <- residuals(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_1_dataset)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Contrasts:
```{r}
contrasts <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust = "none")$contrasts)

# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response")$emmeans) 

## Make a data-set with results:
results <- as.data.frame(contrasts)
results <- results %>%
  separate(contrast, into = c("group1", "group2"), sep = " / ")

results

contrast_rep_F7_females <- results

# Visualize the emmeans:
plot(emmeans(model_1, pairwise ~ rep, 
                      type="response"))

```
# Plot - emmeans
```{r}
pallete_treatments <- c("#af8dc3","#af8dc3",
                        "#af8dc3","#af8dc3",
                        "#af8dc3","#af8dc3")

plot3 <- ggplot(predicted_means, aes(x = rep, y = response, fill = rep)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.9),
                width = 0.25) + 
  scale_fill_manual(values = pallete_treatments) +  
  labs(title = "",
       x = "Replicate",
       y = "Dry weight (mg / fly)") + 
  scale_x_discrete(labels = (c("1", "2", "3", "4", "5", "6"))) +
  theme_classic() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 12),
        aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0),
    axis.text.x = element_text(color="black", size = 12),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color="black", size = 12)) +
  coord_cartesian(ylim=c(0, 0.4))

plot3

```
# #F7  Males
# Fit the model
```{r}
model_1 <- lm(log10(mg_fly) ~ rep, data = big_F7_males) 

summary(model_1)

```
# Check model assumptions:
```{r}
#Create model_1_dataset
model_1_dataset <- big_F7_males
## Introduce the residuals in the data-set:
model_1_dataset$residuals <- residuals(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_1_dataset)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Contrasts:
```{r}
contrasts <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust ="none")$contrasts) 


# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust ="tukey")$emmeans) 

contrasts
predicted_means

## Make a data-set with results:

results <- as.data.frame(contrasts)
results <- results %>%
  separate(contrast, into = c("group1", "group2"), sep = " / ")


contrast_rep_F7_males <- results

# Visualize the emmeans:
plot(emmeans(model_1, pairwise ~ rep, 
                      type="response"))

```
# Plot - emmeans
```{r}
pallete_treatments <- c("#af8dc3","#af8dc3",
                        "#af8dc3","#af8dc3",
                        "#af8dc3","#af8dc3")

plot4 <- ggplot(predicted_means, aes(x = rep, y = response, fill = rep)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.9),
                width = 0.25) + 
  scale_fill_manual(values = pallete_treatments) +  
  labs(title = "",
       x = "Replicate",
       y = "Dry weight (mg / fly)") + 
  scale_x_discrete(labels = (c("1", "2", "3", "4", "5", "6"))) +
  theme_classic() +
   theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 12),
        aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0),
    axis.text.x = element_text(color="black", size = 12),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color="black", size = 12)) + 
  coord_cartesian(ylim=c(0, 0.25))

plot4

```
# #F31  Females
# Fit the model
```{r}
model_1 <- lm(log10(mg_fly) ~ rep, data = big_F31_females) 
summary(model_1)

```
# Check model assumptions:
```{r}
#Create model_1_dataset
model_1_dataset <- big_F31_females
## Introduce the residuals in the dataset:
model_1_dataset$residuals <- residuals(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_1_dataset)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Remove 1 outlier and fit the model:
```{r}
model_1_dataset$unique_id <- paste(model_1_dataset$CGE, model_1_dataset$rep, model_1_dataset$subrep, sep="_")

model_1_dataset <- model_1_dataset %>% dplyr::filter(!unique_id %in% c("CGE3_2_32_2")) 

model_1 <- lm(log10(mg_fly) ~ rep, data = model_1_dataset) 

summary(model_1)

# Check model assumptions:

## Introduce the residuals in the dataset:
model_1_dataset$residuals <- residuals(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_1_dataset)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Contrasts:
```{r}
contrasts <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response",
                      adjust="none")$contrasts) 

# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ rep,
                                   type="response",
                      adjust="tukey")$emmeans) 

## Make a data-set with results:

results <- as.data.frame(contrasts)
results <- results %>%
  separate(contrast, into = c("group1", "group2"), sep = " / ")

results

contrast_rep_F31_females <- results

# Visualize the emmeans:
plot(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="tukey"))

```
# Plot - emmeans
```{r}
pallete_treatments <- c("#762a83","#762a83",
                        "#762a83","#762a83",
                        "#762a83","#762a83")

plot5 <- ggplot(predicted_means, aes(x = rep, y = response, fill = rep)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.9),
                width = 0.25) + 
  scale_fill_manual(values = pallete_treatments) +  
  labs(title = "",
       x = "Replicate",
       y = "Dry weight (mg / fly)") + 
  scale_x_discrete(labels = (c("1", "2", "3", "4", "5", "6"))) +
  theme_classic() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 12),
        aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0),
    axis.text.x = element_text(color="black", size = 12),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color="black", size = 12)) +
  coord_cartesian(ylim=c(0, 0.4))


plot5

```
# #F31  Males
# Fit the model
```{r}
model_1 <- lm(log10(mg_fly) ~ rep , data = big_F31_males) 
summary(model_1)

```
# Check model assumptions:
```{r}
#Create model_1_dataset
model_1_dataset <- big_F31_males
## Introduce the residuals in the dataset:
model_1_dataset$residuals <- residuals(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. # Homogeneous variance! #p-value = 0.3298
leveneTest(residuals ~ rep, data = model_1_dataset)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Remove 2 outliers and fit the model:
To make residuals normal
```{r}
model_1_dataset$unique_id <- paste(model_1_dataset$CGE, model_1_dataset$rep, model_1_dataset$subrep, sep="_")

model_1_dataset <- model_1_dataset %>% dplyr::filter(!unique_id %in% c("CGE3_2_31_2","CGE3_2_32_1")) 

model_1 <-lm(log10(mg_fly) ~ rep, data = model_1_dataset) 
summary(model_1)

```
# Check model assumptions:
```{r}
## Introduce the residuals in the dataset:
model_1_dataset$residuals <- residuals(model_1)

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1))
# Levene's test with one independent variable. 
leveneTest(residuals ~ rep, data = model_1_dataset)
# Q-Q plot of normal residuals
qqPlot(residuals(model_1))

# Normality test1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

```
# Contrasts:
```{r}
contrasts <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="none")$contrasts)


# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="tukey")$emmeans) 

contrasts

predicted_means

## Make a data-set with results:

results <- as.data.frame(contrasts)
results <- results %>%
  separate(contrast, into = c("group1", "group2"), sep = " / ")

results
contrast_rep_F31_males <- results


# Visualize the emmeans:
plot(emmeans(model_1, pairwise ~ rep, 
                      type="response", adjust="tukey"))

```
# Plot - emmean
```{r}
pallete_treatments <- c("#762a83","#762a83",
                        "#762a83","#762a83",
                        "#762a83","#762a83")

plot6 <- ggplot(predicted_means, aes(x = rep, y = response, fill = rep)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.9),
                width = 0.25) + 
  scale_fill_manual(values = pallete_treatments) +  
  labs(title = "",
       x = "Replicate",
       y = "Dry weight (mg / fly)") + 
  scale_x_discrete(labels = (c("1", "2", "3", "4", "5", "6"))) +
  theme_classic() +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 12),
        aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0),
    axis.text.x = element_text(color="black", size = 12),
    axis.ticks = element_line(color = "black"),
    axis.text.y = element_text(color="black", size = 12)) +
  coord_cartesian(ylim=c(0, 0.25))
# Add asterisks manually:
  #annotate("text", x = 2, y = 110, label = "***", size = 6, color = "black") + 
  #annotate("text", x = 5, y = 105, label = "***", size = 6, color = "black")   

plot6

```
# Plot - emmeans females + males
```{r}
combined_plots <- plot3 + plot5 + plot4 + plot6

combined_plots

combined_plot <- plot3 + plot5 + plot4 + plot6 & 
  theme(axis.title.x = element_blank(),  
    axis.text.x = element_text(size = 14),  
    axis.ticks.x = element_line()) 

combined_plot <- wrap_elements(panel = combined_plot) +
  labs(tag = "Replicate") +
  theme(
    plot.tag = element_text(size = 14),
    plot.tag.position = "bottom")


# Export plot:
#ggsave("output_files/plots/High_order_phenotypes/dry_weight_model_rep_wise.svg", combined_plots)

```
# Final p-values: correction across all contrasts
The p-value of emmeans was not corrected (p.value)
```{r}
contrast_rep_F7_females$group <- "females_F7"
contrast_rep_F7_males$group <- "males_F7"
contrast_rep_F31_females$group <- "females_F31"
contrast_rep_F31_males$group <- "males_F31"

combined_contrasts_weight_rep <- rbind(contrast_rep_F7_females,contrast_rep_F7_males,contrast_rep_F31_females,contrast_rep_F31_males)

# Correct for multiple testing across all contrasts and sex:
combined_contrasts_weight_rep$padj_fdr <- p.adjust(combined_contrasts_weight_rep$p.value, method = "fdr", n = length(combined_contrasts_weight_rep$p.value))
combined_contrasts_weight_rep

# Export results:
#write.xlsx(results, file = "output_files/S4_models_high_order_phenotypes.xlsx", sheetName="weight_repwise", append=TRUE)

```
--- END ---