---
title: "High-Protein BigCages - TAG content quantifications from absorbance"
Author: "Claudia Ramirez-Lanzas"
Date: "4/11/2024"
---

Here we compute TAG concentrations using the raw measurements of the
colorimetric assays (absorbance).
Standard curves based on inverse regression polynomial curve for TAG 
samples are used to compute concentration. 

# Load libraries:
```{r}
# Data handling:
library("readxl")
library("writexl")
library("xlsx")
library("tidyverse")

# Calculations:
library("reshape2")


# Plotting:
library("patchwork")
library("ggplot2")
library("ggpubr") 


```
# Imput files:
```{r}
# Load data-set with  standards for TAG.
Big_standards <- read_csv("input_files/BigCages_TAG_540nm_standards.csv", show_col_types = FALSE)

# Import data-set with samples absorbances:
Big_absorbances <- read_csv("input_files/BigCages_TAG_540nm_absorbances.csv", show_col_types = FALSE)

```
* Big_standards: The data-set contains absorbances at 540nm for the standards
where concentration are known. The measurements were corrected by the sample background ABS.
* Big_absorbances: The data-set contains absorbance at 540nm for the samples.
The measurements were corrected by the sample background ABS.

# Function 1:
Create a data-frame with 2nd order polynomial models fitted by quantification batch.
We fit inverse polynomial regression as suggested in
(Delgado R. 2022 Misuse of Beer–Lambert Law and other calibration curves.
R. Soc. Open Sci. 9: 211103. https://doi.org/10.1098/rsos.211103)
```{r}
models_func <- function(x){
  list_batches = split(x, list(x$batch,x$metabolite))
  result <- NULL
  idx <- 0
  for (i in list_batches){
    idx <- idx + 1
    batch_name <- names(list_batches)[idx]
  
    lm2 = lm(ug_ml_expected ~ poly(ABS, 2, raw = TRUE), data = i)
    coeff2 <- round(lm2$coefficients , 7)
    i$lm2_intercept = coeff2[1]
    i$lm2_slope = coeff2[2]
    i$lm2_x2term = coeff2[3]
    i$lm2_std_res = rstandard(lm2)
    i$lm2_R2_adjusted = round(summary(lm2)$adj.r.squared,3)

    plot_result = plotting_func(i, batch_name)
    
    if (is.null(result)) {
      result <- i
    } else {
      result <- full_join(result, i)
    }
  }
  
  result
}

```
# Function 2:
Plot standards curves with fitting models equations:
```{r}

plotting_func <- function(x, plot_name="default"){
  p <- plot(x$ABS,x$ug_ml_expected, main=plot_name)
  lm1 = lm(ug_ml_expected ~ ABS, data = x) # Inverse regression, concentration as a function of known absorbance. 
  lm2 = lm(ug_ml_expected ~ poly(ABS, 2, raw = TRUE), data = x)

  # Plot the predicted lines:
  #p <- p + abline(lm1, col = "green") # 1st degree
  lm2sample <- seq(0,3, length=50)
  p <- p + lines(lm2sample, predict(lm2, data.frame(ABS=lm2sample)), col="red") #2nd degree
  # Get model coefficients:
  coeff1 <- round(lm1$coefficients , 7)
  coeff2 <- round(lm2$coefficients , 7)
  # Plot the ecuations:
  #p <- p + text(0.3, 1500 , paste("lm1: ", coeff1[1], " + ", coeff1[2],
  #                              "*x", "R2 adjusted: ", round(summary(lm1)$adj.r.squared, 3) ), cex=0.7, col="green")
  p <- p + text(0.05, 20 , paste("lm2: ", coeff2[1], " + ", coeff2[2],
                                 "*x", "+", coeff2[3], "*x^2", "R2 adjusted: ",
                                 round(summary(lm2)$adj.r.squared, 3) ), cex=0.7, col="red")
  
  return(p)
}

```
# Function 3:
Computes the coefficient of variation between 2 samples
NOTE: standard deviation is computed for 2 degrees of freedom 
```{r}

my_CV_2df <- function(v1, v2){
  m = mean(c(v1,v2))
  summa_diff_squared = (abs(v1-m) + abs(v2-m))**2
  sq_sdif = sqrt(summa_diff_squared)
  my_sd = sq_sdif/2
  my_cv = (my_sd/m)*100
  return(my_cv)
}

```

# ####Part 1: create standard curves for each quantification batch
# Compute average absorbance:
```{r}
Big_standards$average_ABS <- (Big_standards$ABS1 + Big_standards$ABS2)/2

# Set variables types:
cols_factor <- c( "standard_group","metabolite","batch","point")
cols_numeric <- c("ug_ml_expected", "ABS1","ABS2", "average_ABS")
Big_standards[cols_factor] <- lapply(Big_standards[cols_factor], as.factor)
Big_standards[cols_numeric] <- lapply(Big_standards[cols_numeric], as.numeric)

str(Big_standards)
# Convert data-set into long format:
Big_standards <- pivot_longer(Big_standards, cols = c("ABS1","ABS2"),
                            names_to = "tech_rep", values_to = "ABS")
Big_standards$ABS <- as.numeric(Big_standards$ABS)
Big_standards$tech_rep <- as.factor(Big_standards$tech_rep)


Big_standards

```
# Run models_func to fit regression models:
```{r}
dataset_TAG_residuals <- models_func(Big_standards)

```
# Filter outlier standardized residuals and run the standard curves again:
In order to automate the optimization of standard curves we remove the
outlier standardized residuals of each standard curve.

Note: A standardized residual is the raw residual divided by an estimate 
of the standard deviation of the residuals. 
There is a general consensus in literature that standardized residuals (absolute value) 
larger than 2 can be outliers.

```{r}
# Use polynomial model. 

dataset_TAG_residuals_filter_lm2 <- dataset_TAG_residuals %>% filter(abs(lm2_std_res) < 2)  
bc <- nrow(dataset_TAG_residuals) - nrow(dataset_TAG_residuals_filter_lm2) # 3 residuals removed

print("residuals removed:")
bc

```
# Run models_func to fit models AFTER FILTERING:
```{r}
dataset_TAG_residuals_filter_lm2_refit <- models_func(dataset_TAG_residuals_filter_lm2) 

```
# Visualize model coefficients in a table:
```{r}
# Remove all duplicated rows:
TAG_dataset_models <- distinct(dataset_TAG_residuals_filter_lm2_refit, standard_group, 
                               metabolite, batch, 
                               .keep_all= TRUE)

TAG_dataset_models

```
# ####Part 2:  Compute concentration (ug/ml) from absorbances:
# Format variables:
```{r}
# Set variables types:
cols_factor <- c( "rep", "subrep",
                  "treatment","unique_id",
                  "id_quantification", "sample",
                  "tech_rep", "sex", "standard_group", "Who")

cols_numeric <- c("nb_flies", "total_g",
                  "total_mg", "dilution_factor",
                  "ABS", "hom_vol_ml")

Big_absorbances[cols_factor] <- lapply(Big_absorbances[cols_factor], as.factor)
Big_absorbances[cols_numeric] <- lapply(Big_absorbances[cols_numeric], as.numeric)


(str(Big_absorbances))

```

# Compute concentrations:
```{r}
# First: merge dataset_TAG with TAG_dataset_models so each sample (row) 
#contains also columns for the coefficient of the metabolite-batch specific models:

merge_1 <- Big_absorbances
merge_2 <- TAG_dataset_models %>% select(-c(metabolite, batch, point, ug_ml_expected, average_ABS, tech_rep, ABS, Who))

merged_dataset_samples_and_models <- inner_join(merge_1, merge_2, by="standard_group")
merged_dataset_samples_and_models <- merged_dataset_samples_and_models %>% relocate(ABS, .after = lm2_R2_adjusted)

# Second: compute homogenate concentration (ug_ml) in a new column by using next formula:
# (concentration = (absorbance x slope) + intercept):
merged_dataset_samples_and_models <- merged_dataset_samples_and_models %>%  
  rowwise() %>% 
  mutate(conc_ug_ml = (lm2_intercept + ABS*lm2_slope + lm2_x2term*(ABS)^2)) # Using polynomial curve. 

merged_dataset_samples_and_models$conc_ug_ml <- as.numeric(merged_dataset_samples_and_models$conc_ug_ml)

```

# Compute CV% (coefficient of variance): - tech reps
For each pair of technical replicates we compute the coefficient of variance in order to 
filter samples above a threshold.
```{r}
# First spread tech_rep_nb column in 4 columns so CV can be computed row-wise: 

short_df <- merged_dataset_samples_and_models %>% dplyr::select("unique_id",
                                                                "sample", "rep", "subrep",
                                                                "treatment","nb_flies", "hom_vol_ml",
                                                                "sex", "total_g", "total_mg",
                                                 "dilution_factor",
                                                 "standard_group",
                                                 "tech_rep",
                                                 "conc_ug_ml",
                                                 "Who")

short_wide_df <- as.data.frame(short_df) %>% dcast(sample +  
                                                     treatment + sex +
                                                     nb_flies +
                                                     hom_vol_ml +
                                                     total_g +
                                                     total_mg + 
                                                     dilution_factor ~ tech_rep, value.var = "conc_ug_ml")


# Format variables:
cols_numeric <- c("tech_rep_1", "tech_rep_2", "tech_rep_3", "tech_rep_4")
short_wide_df[cols_numeric] <- lapply(short_wide_df[cols_numeric], as.numeric)

#Compute CV%
short_wide_df <- short_wide_df %>% as_tibble() %>% rowwise() %>% 
  mutate(CV_concentration_1and2 = my_CV_2df(tech_rep_1, tech_rep_2),
         CV_concentration_3and4 = my_CV_2df(tech_rep_3, tech_rep_4),
         CV_concentration_1and3 = my_CV_2df(tech_rep_1, tech_rep_3),
         CV_concentration_1and4 = my_CV_2df(tech_rep_1, tech_rep_4),
        CV_concentration_2and3 = my_CV_2df(tech_rep_2, tech_rep_3),
         CV_concentration_2and4 = my_CV_2df(tech_rep_2, tech_rep_4))

```
# Check repeated samples:
Some samples were repeated and have several measurements. Keep those pairs with lowest CV.
```{r}

# prepare data-set:
short_df$full_unique_id <- paste(short_df$unique_id, short_df$tech_rep, sep ="_")

# Manually identify good samples:
# CGE2_2_36_2_male tech reps 2 & 3
# CGE1_B_8_male 3 4
# CGE1_B_7_female 3 4


# Samples to be removed: 
sample_list <- c("tag_batch_1_CGE2_2_36_2_male_tech_rep_1",
                 "tag_batch_1_CGE1_B_8_male_tech_rep_1",
                 "tag_batch_1_CGE1_B_8_male_tech_rep_2",
                 "tag_batch_1_CGE1_B_7_female_tech_rep_1",
                 "tag_batch_1_CGE1_B_7_female_tech_rep_2")

# Remove samples:
short_filter_df <- subset(short_df, !(full_unique_id %in% sample_list))
 
```
# Recompute CV:
```{r}

short_wide_df_2 <- as.data.frame(short_filter_df) %>% dcast(sample + rep +
                                                              subrep +
                                                              treatment + sex +
                                                              nb_flies + hom_vol_ml +
                                                              total_g + total_mg +
                                                              dilution_factor +
                                                              Who  ~ tech_rep,
                                                            value.var = "conc_ug_ml")

short_wide_df_2 <- short_wide_df_2 %>%
  mutate(
    tech_rep_1 = coalesce(tech_rep_1, tech_rep_3, tech_rep_4, tech_rep_2),
    tech_rep_2 = coalesce(tech_rep_2, tech_rep_4, tech_rep_3, tech_rep_1)) 

cols_numeric <- c("tech_rep_1", "tech_rep_2")
short_wide_df_2[cols_numeric] <- lapply(short_wide_df_2[cols_numeric], as.numeric)


#### CV%
short_wide_df_2 <- short_wide_df_2 %>% as_tibble() %>% rowwise() %>% 
  mutate(CV_concentration_1and2 = my_CV_2df(tech_rep_1, tech_rep_2))

short_wide_df_2 <- short_wide_df_2 %>% select("sample", "rep", "subrep", "treatment",
                                              "sex", "Who","nb_flies",
                                              "hom_vol_ml", "total_g",
                                              "total_mg", "dilution_factor",
                                              "tech_rep_1", "tech_rep_2",
                                              "CV_concentration_1and2" )

short_wide_df_2 <- short_wide_df_2 %>% as_tibble() %>% rowwise() %>% 
  mutate(mean_conc_ug_ml = mean(c(tech_rep_1, tech_rep_2)))

```
# Histogram of CV:
```{r}
hist(short_wide_df_2$CV_concentration_1and2)
```
# Compute final concentrations: 
```{r}
short_wide_df_2 <- short_wide_df_2 %>% as_tibble() %>% rowwise() %>% 
  mutate(ug_TAG_homogenate = mean_conc_ug_ml* hom_vol_ml)

short_wide_df_2 <- short_wide_df_2 %>% as_tibble() %>% rowwise() %>% 
  mutate(ug_TAG_fly = ug_TAG_homogenate/nb_flies)

short_wide_df_2 <- short_wide_df_2 %>% as_tibble() %>% rowwise() %>% 
  mutate(ug_TAG_mg_fly= ug_TAG_homogenate/total_mg)

```
# Include variables:
```{r}
# Include CGE:
final_dataset <- short_wide_df_2 %>% 
     mutate(CGE = case_when(
       grepl("CGE1_", sample) ~ "CGE1",
       grepl("CGE2_1_", sample) ~"CGE2_1",
       grepl("CGE2_2_", sample) ~"CGE2_2",
       grepl("CGE3_1_", sample) ~"CGE3_1",
       grepl("CGE3_2_", sample) ~"CGE3_2"))

final_dataset <- final_dataset %>% dplyr::relocate(CGE, .after = sample)
final_dataset$CGE <- as.factor(final_dataset$CGE )

# Include generation:
final_dataset <- final_dataset %>% 
     mutate(generation = case_when(
       grepl("CGE1_", sample) ~ "gen0",
       grepl("CGE2_1_", sample) ~ "gen7",
       grepl("CGE2_2_", sample) ~ "gen7",
       grepl("CGE3_1_", sample) ~ "gen31",
       grepl("CGE3_2_", sample) ~ "gen31"))

final_dataset <- final_dataset %>% dplyr::relocate(generation, .after = CGE)
final_dataset$generation <- as.factor(final_dataset$generation )

```

# Remove samples with CV >= 10%
```{r}
final_dataset <- final_dataset %>% dplyr::filter(CV_concentration_1and2 < 10)
  
hist(final_dataset$CV_concentration_1and2)

```
# Export dataset:
```{r}
export_dataset <- final_dataset 
#write_csv(export_dataset, "/Users/cramirez/Desktop/github_repositories/BigCages_high_protein/input_files/BigCagesTAG_concentrations.csv")

```

--- END ---