---
title: "High-Protein BigCages - TAG content"
Author: "Claudia Ramirez-Lanzas"
Date: "4/11/2024"
---

# Load libraries:
```{r}
# Data handling:
library("readxl")
library("writexl")
library("xlsx")
library("tidyverse")
# Stasistics:
library("lme4")
library("lmerTest") # to include approximate p-values in the output of lmer. 
library("emmeans")
library("car")
library("HLMdiag")
library("ggeffects")
# Plotting:
library("patchwork")
library("ggplot2")
library("ggpubr") 


```
# ####Part 1: Investigate raw data
# Import data-set:
```{r}
TAG_Big <- read_csv("input_files/BigCagesTAG_concentrations.csv", show_col_types = FALSE)

```
*samples_dataset: contains TAG concentrations computed using inverse polynomial
regression in 03_TAG_quantification notebook.  

# Prepare data-set:
```{r}
# Relevel generation:
TAG_Big$generation <- factor(TAG_Big$generation, 
                                     levels = c("gen0" , "gen7", "gen31"))

# Include population_time:
TAG_Big$population_time <-  paste(TAG_Big$treatment, 
                                          TAG_Big$generation, sep="_")

# Relabel replicate:
TAG_Big <- TAG_Big %>% mutate(rep = case_when(
  grepl("CGE1", CGE)  ~"base",
  grepl("31_", sample)  ~"01",
  grepl("32_", sample)  ~"02",
  grepl("33_", sample)  ~"03",
  grepl("34_", sample)  ~"04",
  grepl("35_", sample)  ~"05",
  grepl("36_", sample)  ~"06"))

# Include wet mg per fly:
TAG_Big$mg_fly <- TAG_Big$total_mg /  TAG_Big$nb_flies
TAG_Big$mg_fly <- as.numeric(TAG_Big$mg_fly)

TAG_Big <- TAG_Big %>% relocate(rep, .after = CGE)
TAG_Big <- TAG_Big %>% relocate(subrep, .after = rep)
TAG_Big <- TAG_Big %>% relocate(population_time, .after = subrep)

# Include unique id:
TAG_Big$unique_id <- paste(TAG_Big$CGE,
                                 TAG_Big$rep,
                                 TAG_Big$subrep, sep="_")

# Format variables:
TAG_Big$population_time <- as.factor(TAG_Big$population_time)
TAG_Big$sex <- as.factor(TAG_Big$sex)
TAG_Big$treatment <- as.factor(TAG_Big$treatment)
TAG_Big$Who <- as.factor(TAG_Big$Who)
TAG_Big$CGE <- as.factor(TAG_Big$CGE)
TAG_Big$rep <- as.factor(TAG_Big$rep)
TAG_Big$subrep <- as.factor(TAG_Big$subrep)

# Relevel generation:
TAG_Big$generation <- factor(TAG_Big$generation, levels = c("gen0", "gen7", "gen31"))

str(TAG_Big)

```

# Histogram CV:
CV shows the technical variation, as standard conventions it should not exceed 10%.
All samples have <10% CV.

```{r}
hist(TAG_Big$CV_concentration_1and2)

```

## Plot - TAG / fly
```{r}
# ug metabolite per fly per generation:
plot_df <- TAG_Big
# Change generation scale to reduce distances proportionally:
plot_df <- plot_df %>% mutate(gen = case_when(generation == "gen0" ~ 0,
                                              generation == "gen7" ~ 0.5,
                                              generation == "gen31" ~ 1.5))

pallete_treatments <- c("orange","#af8dc3","#762a83")

plot1 <- plot_df %>% 
  ggplot(aes(x = gen, y = ug_TAG_fly, fill = as.factor(gen))) +
  geom_boxplot(outlier.size = 0.3) + 
  labs(title = "") +
  ylab("TAG equivalents (\u00b5g / fly)") +
  xlab("Generation") +
  #scale_x_discrete(labels = c("0", "7", "31")) +  
  scale_fill_manual(values = pallete_treatments) + 
  theme(text = element_text(size = 13)) + 
  facet_wrap(~sex , ncol = 2) +
  theme_classic() +
  theme(aspect.ratio = 1, plot.margin = margin(0, 0, 0, 0)) + 
  theme(axis.text.x.top  = element_blank(),
    axis.line.x.top = element_blank(),
    axis.ticks.x.top = element_blank(),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="none",
    strip.background = element_rect(
     color="white", fill="white")) +
  scale_x_continuous(breaks = c(0, 0.5, 1.5),
                     labels = c(0,7,31))

plot1
plot_df <- TAG_Big

plot_df %>% ggplot(aes(x = as.numeric(generation), y = ug_TAG_fly, color = rep)) +
  geom_point(size=4, alpha=0.4) + facet_wrap(~ sex)
  
```


## Plot - TAG / mg
```{r}
pallete_treatments <-  c("orange","#af8dc3","#762a83")
plot2 <- TAG_Big %>% 
  ggplot(aes(x = generation, y = ug_TAG_mg_fly, fill = generation)) +
  geom_boxplot(outlier.size=0.3) + 
  geom_point(position = position_dodge(width=0.75), 
             aes(group = treatment), size = 0.3) +
  labs(title = "") + ylab("ug TAG per mg") +
  scale_x_discrete(labels = c("0", "7", "31")) +  
  scale_fill_manual(values = pallete_treatments) + 
  theme_light() + 
  theme(text = element_text(size = 13)) + 
  facet_wrap(~sex , ncol = 2) +
  theme_light() + theme(legend.position = "none")

plot2

```

## Plot - Wet weight
```{r}

wet_weight_df <- TAG_Big %>% 
  mutate(mg_fly = total_mg / nb_flies)

pallete_treatments <-  c("orange","#af8dc3","#762a83")

plot3 <- wet_weight_df %>%
  ggplot(aes(x = generation, y = mg_fly, fill = generation)) +
  geom_boxplot(outlier.size=0.3) + 
  labs(title = "") +   geom_point(position = position_dodge(width=0.75), 
             aes(group = treatment), size = 0.3) +
  scale_fill_manual(values = pallete_treatments) + 
  ylab("Wet weight") +   
  scale_x_discrete(labels = c("0", "7", "31")) +  
  scale_fill_manual(values = pallete_treatments) + 
  theme_light() + 
  theme(text = element_text(size = 13)) + 
  facet_wrap(~sex, ncol = 2) +
  theme_light() + theme(legend.position = "none")


plot3


```
# ####Part 2: Overall model 
# Aggregate by mean:
After looking at the raw data, Marlies suggested to aggregate the sub-replicates
using the mean and fitting a simpler model where random slopes are not needed. 
```{r}
model_dataset_full <- TAG_Big %>% dplyr::select(c("unique_id", "CGE", "generation",
                                                        "rep", "subrep",
                                                        "sample", "treatment",
                                                        "Who",
                                                        "sex", "nb_flies",
                                                        "total_mg", "mg_fly",
                                                 "ug_TAG_mg_fly",
                                                 "ug_TAG_fly"))


# The samples from base population are not averaged:
base <- model_dataset_full %>% dplyr::filter(CGE == "CGE1")

# The samples from evolved populations are averaged within each replicate:
evolved <- model_dataset_full %>% dplyr::filter(CGE != "CGE1")
evolved_aggre <- evolved %>% 
  group_by(generation, CGE,rep, sex) %>% 
  summarise(ug_TAG_fly = mean(ug_TAG_fly),
            mg_fly = mean(mg_fly))

# Select columns to match with evolved_aggre:
base <- base %>% dplyr::select(c(colnames(evolved_aggre)))

# Combine base and aggregated evolved replicates:
model_dataset_aggre <- rbind(base, evolved_aggre)

# Format the variables:
cols_factor <- c("rep", 
                 "generation",
                 "sex")

model_dataset_aggre[cols_factor] <- lapply(model_dataset_aggre[cols_factor], as.factor)

table(model_dataset_aggre$rep, model_dataset_aggre$generation)
  
```
# Fit the model 
Suggested by Marlies
```{r}
model_1 <- lmer(log10(ug_TAG_fly) ~ generation*sex + mg_fly + (1 | rep), data = model_dataset_aggre, REML = FALSE)

summary(model_1)

# Introduce the residuals in the data-set:
model_dataset_1 <- model_dataset_aggre
model_dataset_1$residuals <- residuals(model_1)
model_dataset_1$unique_id <- paste(model_dataset_1$CGE, model_dataset_1$rep, model_dataset_1$sex, sep="_")

```
## Test model assumptions:
```{r}

# Homogeneity of variance:
plot(predict(model_1), residuals(model_1)) # Predicted values are plotted against residuals to assess variance homogeneity.
leveneTest(residuals ~ generation, data = model_dataset_1) # Levene's test with one independent variable. 

# Normality of residuals:
qqPlot(residuals(model_1)) # Q-Q plot of normal residuals.

# Normality test 1:
shapiro.test(residuals(model_1)) 

# Normality test 2:
#Kolmogorov-Smirnov test
ks.test(residuals(model_1),"pnorm", mean(residuals(model_1)), sd(residuals(model_1))) 

# Check assumptions for random intercept term:
# Are random intercept normally distributed?
ranef_rep <- ranef(model_1)[[1]]
ranef_rep$rep <- rownames(ranef_rep)
ranef_rep <- merge(ranef_rep, model_dataset_1[!duplicated(model_dataset_1$rep), 
                                              c("generation", "rep")], by = "rep")

qqnorm(ranef_rep$`(Intercept)`, 
       main="QQ Plot for random intercepts per generation", 
       col = ranef_rep$generation, pch =19)


```
# Contrasts:
```{r}
# Contrasts:
contrasts <- summary(emmeans(model_1, pairwise ~ generation|sex, 
                      type="response", adjust = "none")$contrasts) 
contrasts

# Predicted Means:
predicted_means <- summary(emmeans(model_1, pairwise ~ generation|sex, 
                      type="response", adjust="none")$emmeans) 

# Change generation scale to reduce distances proportionally: (make nicer plot)
predicted_means <- predicted_means %>% dplyr::mutate(generation = case_when(generation == "gen0" ~ 0, 
                                                                            generation == "gen7" ~ 0.5,
                                                                            generation == "gen31" ~ 1.5))

# Plot:
plot(emmeans(model_1, pairwise ~ generation|sex, 
                      type="response", adjust = "none"))


## Make a data-set with results:
results <- as.data.frame(contrasts)
results <- results %>%
  separate(contrast,
           into = c("group1", "group2"), sep = " / ")

results


```
# Final p-values: correction across all contrasts
The p-value of emmeans was not corrected (p.value).
Here we use FDR correction, suggested by Marlies.
```{r}
# Correct for multiple testing across all contrasts and sex:
results$padj_fdr <- p.adjust(results$p.value, method = "fdr",
                               n = length(results$p.value))
results

# Export results:
#write.xlsx(results, file = "output_files/S4_models_high_order_phenotypes.xlsx", sheetName="TAG_overall", append=TRUE)

```
## Plot model results: 
```{r}

pallete_treatments <- c("#f1a340","#af8dc3", "#762a83")

plot_overall_tag <- predicted_means %>% ggplot(
  aes(x = generation,
      y = response, fill=as.factor(generation))) + 
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9),
           width=0.3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.7),
                width = 0.1) +
  facet_wrap(~sex, 
             labeller = as_labeller(
               c(female = "Females", male = "Males")), scales = "free_y") +
  scale_fill_manual(values = pallete_treatments) +  
  labs(title = "",
       x = "",
       y = "TAG equivalents (ug / fly)") +  
  theme_classic() +
  xlab("Generation") + 
  theme(aspect.ratio = 1, plot.margin = margin(0, 0, 0, 0)) + 
  theme(axis.text.x.top  = element_blank(),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    strip.background = element_blank(),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 16, face = "bold"),
    legend.position="none") +
  scale_x_continuous(breaks = c(0, 0.5, 1.5),
                     labels = c(0,7,31))


plot_overall_tag

# Export plot:
#ggsave("output_files/plots/High_order_phenotypes/TAG_overall_emmeans.svg",
#plot_overall_fec, units = "mm", height = 150, width = 150)

```

--- END ---