---
title: "High-Protein BigCages - compute gene expression variance"
Author: "Claudia Ramirez-lanzas"
Date: "27/05/2024"
---

# Load libraries:
```{r}
# Plotting:
library("readxl") 
library("writexl") 
library("tidyverse") 

# Plotting
library("khroma") # plotting colour blind safe
library("ggplot2") 
library("patchwork") # to combine plots
library("svglite") 
library("egg") #
library("RColorBrewer") 
library("ggbreak")

```
# Set-up colors:
```{r}
bright <- color("bright")
plot_scheme(bright(6), colours = TRUE, names = TRUE, size = 0.9)
#
muted <- color("muted")
plot_scheme(muted(9), colours = TRUE, names = TRUE, size = 0.9)

color_plateau <- muted(9)[2] #"#332288"
color_monotonic <- muted(9)[7] #"#44AA99"
color_late_response <-  "#D4AF37"
color_incomplete_reversed <- muted(9)[9] #"#AA4499"
color_complete_reversed <- muted(9)[5] #"#88CCEE" 
color_shared <- muted(9)[1] #"#CC6677" 

plot_scheme(muted(9)[2], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[7], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[9], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[3], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[5], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[1], colours = TRUE, names = TRUE, size = 0.9)


```
# Import files:
```{r}
classify_df <- read_csv("input_files/classify_trajectories_df.csv", show_col_types = FALSE)

count_table <- read.delim("input_files/count_table.csv", sep=";")
count_table <- count_table %>% column_to_rownames(var = "gene_id")

```

*classify_df: contains all genes classified by transcriptomic
trajectory along with the results of DE analysis. This dataset was generated in
08_Transcriptomic_trajectories notebook.

*count_table: .csv file  containing the RNA count table (the output of
featureCounts function from subRead).

# # Part 1: compute variance - per gene 
Here we compute the variance of each gene´s expression across replicates
within each generation.
# Estimate library size:
Library size = Total number of fragments (paired ends) mapped to the genome. 
```{r}
lib_size <- colSums(count_table)
hist(lib_size)
average_library_size <- mean(lib_size) # 10 millions
average_library_size

```
### Remove lowly expressed genes:
With an average library size of around 10M, a CPM =  1 corresponds to 10 counts 
per library.
We remove genes with <10 counts per library in 1 or more samples.

Following criteria in:
Chen Y, Lun ATL and Smyth GK. From reads to genes to pathways:
differential expression analysis of RNA-Seq experiments using 
Rsubread and the edgeR quasi-likelihood pipeline [version 2; 
peer review: 5 approved]. F1000Research 2016, 5:1438
(https://doi.org/10.12688/f1000research.8987.2)

```{r}
#Set CPM threshold:
CPM_threshold <- 1 

# Create filtered count table:
count_table_filt <- count_table[apply(cpm(count_table), 1, function(x){!(sum(x < CPM_threshold) >= 1)}),] # If there is one sample or more with CPM < threshold then the gene is filtered out. 

# print filtering summary:
paste("all genes:", nrow(count_table))
paste("genes available after filtering:", nrow(count_table_filt))
paste("genes discarded", nrow(count_table) - nrow(count_table_filt))
paste("% genes discarded" , ((nrow(count_table) - nrow(count_table_filt))/nrow(count_table))*100)

```
### Prepare the data-set
```{r}
gene_per_group <- classify_df %>% dplyr::select(gene_id, group, direction, full_group)

# Get the gene_id of all genes: (after filtering low expression)
all_genes_id <- gene_per_group %>% dplyr::select("gene_id")

# Get count table after filtering low expression:
count_all_genes <- count_table_filt 

# Compute log2CPM: (log = FALSE)
log2cpm_all_genes <- edgeR::cpm(as.matrix(count_all_genes), 
       normalized.lib.sizes = TRUE,
       log = TRUE)

# Gene_id to column:
count_all_genes <- rownames_to_column(as.data.frame(count_all_genes), var = "gene_id") 

# Include group information: 
log2cpm_all_genes <- rownames_to_column(as.data.frame(log2cpm_all_genes), var = "gene_id") 
log2cpm_all_genes_group <- left_join(log2cpm_all_genes, gene_per_group, by = "gene_id")
log2cpm_all_genes_group <- relocate(log2cpm_all_genes_group, group, .after = gene_id)
log2cpm_all_genes_group <- relocate(log2cpm_all_genes_group, direction, .after = group)
log2cpm_all_genes_group <- relocate(log2cpm_all_genes_group, full_group, .after = direction)

# Order column names (samples) alphabetically: 
log2cpm_all_genes_group <- log2cpm_all_genes_group %>% dplyr::select(order(colnames(log2cpm_all_genes_group)))
log2cpm_all_genes_group <- log2cpm_all_genes_group %>% dplyr::relocate(gene_id, .before = CGE1_rep_B_1)
log2cpm_all_genes_group <- log2cpm_all_genes_group %>% dplyr::relocate(group, .before = CGE1_rep_B_1)
log2cpm_all_genes_group <- log2cpm_all_genes_group %>% dplyr::relocate(direction, .before = CGE1_rep_B_1)
log2cpm_all_genes_group <- log2cpm_all_genes_group %>% dplyr::relocate(full_group, .before = CGE1_rep_B_1)

```
## Compute the mean CPM of sub-replicates:
The expression of the sub-replicates of each evolved replicate
are averaged (except for base population).
```{r}

# Compute the mean CPM of sub-replicates within each replicate (except Base):
rep_aggragated_mean_log2CPM <- log2cpm_all_genes_group[, 1:9]

# Calculate the mean of each replicate:
rep_aggragated_mean_log2CPM$CGE2_1_rep_31 <- apply(log2cpm_all_genes_group[, 10:12], 1, mean)
rep_aggragated_mean_log2CPM$CGE2_1_rep_32 <- apply(log2cpm_all_genes_group[, 13:15], 1, mean)
rep_aggragated_mean_log2CPM$CGE2_1_rep_33 <- apply(log2cpm_all_genes_group[, 16:18], 1, mean)
rep_aggragated_mean_log2CPM$CGE2_1_rep_34 <- apply(log2cpm_all_genes_group[, 19:21], 1, mean)
rep_aggragated_mean_log2CPM$CGE2_1_rep_35 <- apply(log2cpm_all_genes_group[, 22:24], 1, mean)
rep_aggragated_mean_log2CPM$CGE2_1_rep_36 <- apply(log2cpm_all_genes_group[, 25:27], 1, mean)
rep_aggragated_mean_log2CPM$CGE3_1_rep_31 <- apply(log2cpm_all_genes_group[, 28:30], 1, mean)
rep_aggragated_mean_log2CPM$CGE3_1_rep_32 <- apply(log2cpm_all_genes_group[, 31:33], 1, mean)
rep_aggragated_mean_log2CPM$CGE3_1_rep_33 <- apply(log2cpm_all_genes_group[, 34:36], 1, mean)
rep_aggragated_mean_log2CPM$CGE3_1_rep_34 <- apply(log2cpm_all_genes_group[, 37:39], 1, mean)
rep_aggragated_mean_log2CPM$CGE3_1_rep_35 <- apply(log2cpm_all_genes_group[, 40:42], 1, mean)
rep_aggragated_mean_log2CPM$CGE3_1_rep_36 <- apply(log2cpm_all_genes_group[, 43:45], 1, mean)

# Make long format:
log2CPM_sample_group_long <- pivot_longer(cols = -c("gene_id","group", "direction", "full_group"), 
                                           data = rep_aggragated_mean_log2CPM, 
                                           names_to = "sample", 
                                           values_to = "log2CPM")

# Include generation:
log2CPM_sample_group_long <- log2CPM_sample_group_long %>% 
  mutate(generation = case_when(grepl("CGE1", sample) ~ "gen0",
                           grepl("CGE2", sample) ~ "gen07",
                           grepl("CGE3", sample) ~ "gen31"))

# Create generation_genID column:
log2CPM_sample_group_long$generation_genID <- paste(log2CPM_sample_group_long$generation,
                                                log2CPM_sample_group_long$gene_id,
                                                sep ="_")

# Keep needed columns:
gene_expression_df <- log2CPM_sample_group_long %>% 
  dplyr::select(c("group", "direction", "full_group", "generation", "sample",
                  "gene_id", "generation_genID", "log2CPM"))

# Convert it into data-frame:
gene_expression_df <- as.data.frame(gene_expression_df)

# Format variables:
gene_expression_df$generation <- as.factor(gene_expression_df$generation)
gene_expression_df$group <- as.factor(gene_expression_df$group)
gene_expression_df$direction <- as.factor(gene_expression_df$direction)
gene_expression_df$full_group <- as.factor(gene_expression_df$full_group)
gene_expression_df$log2CPM <- as.numeric(gene_expression_df$log2CPM)

```
## Computer variance per gene:
```{r}
variance_per_gene <- gene_expression_df %>%
  dplyr::filter(!group %in% c("transitory_F7_F31"))%>%
  group_by(generation_genID) %>%
  summarize(variance_log2CPM = var(log2CPM),
            .groups = 'drop')

# Include grouping variables:
variance_per_gene_df <- gene_expression_df %>% 
  dplyr::filter(!group %in% c("transitory_F7_F31")) %>%
  left_join(variance_per_gene, by = c("generation_genID"))

variance_per_gene_df$group <- factor(variance_per_gene_df$group,
                           levels = c("plateau",
                                      "monotonic",
                                      "late_response",
                                      "incomplete_reversed",
                                      "complete_reversed"))

# Calculate median/mean and confidence intervals:
summary_data <- variance_per_gene_df %>%
  group_by(generation,
           direction,
           group,
           full_group) %>%
  summarise(
    median_log2CPM_variance = median(variance_log2CPM),
    mean_log2CPM_variance = mean(variance_log2CPM),
    ci_lower = quantile(variance_log2CPM, 0.025),
    ci_upper = quantile(variance_log2CPM, 0.975),
    .groups = 'drop')

summary_data

# Calculate median/mean and confidence intervals:
summary_data_2 <- variance_per_gene_df %>%
  group_by(generation,
           group) %>%
  summarise(
    median_log2CPM_variance = median(variance_log2CPM),
    mean_log2CPM_variance = mean(variance_log2CPM),
    ci_lower = quantile(variance_log2CPM, 0.025),
    ci_upper = quantile(variance_log2CPM, 0.975),
    .groups = 'drop')

# Remove non-significant genes:
summary_data_2 <- summary_data_2 %>% drop_na(group)

```
# Export Variance per gene dataset:
```{r}
export_df <- variance_per_gene_df %>% droplevels()
#write_csv(export_df, "output_files/variance_per_gene.csv")

```
## Lineplots: - medians across genes
Before, we computed variance for each gene. Here we plot the medians of variances 
of each transcriptomic group.
```{r}

plot_df <- summary_data %>% dplyr::filter(full_group != "non_significant_none")

plot_df %>%
  ggplot(aes(x = generation,
             y = median_log2CPM_variance,
             color = group)) +
  geom_point(shape=1, size=2) + 
  geom_line(aes(group = group), 
            alpha = 1, linewidth =1) +  
  ylab(bquote(Median~of~gene~xpression~(log[2]*CPM)~variance)) +
  xlab("Generation") +
  scale_color_manual(values =  c(color_plateau,
                                color_monotonic,
                                color_late_response,
                               color_incomplete_reversed,
                               color_complete_reversed),
                     labels = c("Plateau",
                                "Monotonic",
                                "Late-response",
                               "Incomplete-reverse",
                               "Complete-reverse")) +
  theme_classic() +
  facet_wrap(~ direction, nrow=1, labeller = labeller(direction = 
    c("down" = "down-regulated",
      "up" = "up-regulated"))) +
  theme(
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="black",
                                    fill="white")) +
  scale_x_discrete(labels = c("0", "7", "31"))



##

plot_df <- summary_data_2 %>% dplyr::filter(group != "non_significant_none")

plot_df %>%
  ggplot(aes(x = generation,
             y = median_log2CPM_variance,
             color = group)) +
  geom_point(shape=1, size=2) + 
  geom_line(aes(group = group), 
            alpha = 1, linewidth =1) +  
  ylab(bquote(Median~of~gene~xpression~(log[2]*CPM)~variance)) +
  xlab("Generation") +
  scale_color_manual(values =  c(color_plateau,
                                color_monotonic,
                                color_late_response,
                               color_incomplete_reversed,
                               color_complete_reversed),
                     labels = c("Plateau",
                                "Monotonic",
                                "Late-response",
                               "Incomplete-reverse",
                               "Complete-reverse")) +
  theme_classic() +

  theme(
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="black",
                                    fill="white")) +
  scale_x_discrete(labels = c("0", "7", "31"))

```
## Lineplots: - means across genes
Before, we computed variance for each gene. Here we plot the medians of variances 
of each transcriptomic group.
```{r}

plot_df <- summary_data %>% dplyr::filter(full_group != "non_significant_none")

plot_df %>%
  ggplot(aes(x = generation,
             y = mean_log2CPM_variance,
             color = group)) +
  geom_point(shape=1, size=2) + 
  geom_line(aes(group = group), 
            alpha = 1, linewidth =1) +  
  ylab(bquote(Mean~of~gene~xpression~(log[2]*CPM)~variance)) +
  xlab("Generation") +
  scale_color_manual(values =  c(color_plateau,
                                color_monotonic,
                                color_late_response,
                               color_incomplete_reversed,
                               color_complete_reversed),
                     labels = c("Plateau",
                                "Monotonic",
                                "Late-response",
                               "Incomplete-reverse",
                               "Complete-reverse")) +
  theme_classic() +
  facet_wrap(~ direction, nrow=1, labeller = labeller(direction = 
    c("down" = "down-regulated",
      "up" = "up-regulated"))) +
  theme(
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="black",
                                    fill="white")) +
  scale_x_discrete(labels = c("0", "7", "31"))


#ggplot2::ggsave( "output_files/plots/median_var_gene_plot.svg", last_plot())

```
## Boxplots - variance all genes per generation
```{r}
### DEGs:
plot_df <- variance_per_gene_df 
plot_df <- plot_df %>% mutate(group = case_when(is.na(group) == TRUE ~ "non_significant", TRUE ~ group))

plot_df <- plot_df %>% dplyr::filter(group != "non_significant") %>% droplevels()

ggplot(plot_df, aes(x = generation,
             y = variance_log2CPM)) +
  geom_boxplot() +
  theme_classic() + 
    xlab("Generation") +
  scale_x_discrete(labels = c("0" ,"7", "31")) +
theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 16, color = "black"),
    axis.text.y = element_text(size = 16, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="black",
                                    fill="white")) +
  coord_cartesian(ylim = c(0,0.05))

```
```{r}
### Non DEGs:
plot_df <- variance_per_gene_df %>% mutate(group = case_when(is.na(group) == TRUE ~ "non_significant", TRUE ~ group))
plot_df <- plot_df %>% dplyr::filter(group == "non_significant") %>% droplevels()
  
ggplot(plot_df, aes(x = generation,
             y = variance_log2CPM)) +
  geom_boxplot() +
  theme_classic() + 
    xlab("Generation") +
  scale_x_discrete(labels = c("0" ,"7", "31")) +
theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 16, color = "black"),
    axis.text.y = element_text(size = 16, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="black",
                                    fill="white")) + coord_cartesian(ylim = c(0,0.05)) 



```
## Boxplots all groups:
```{r}
plot_df <- variance_per_gene_df %>%
  mutate(group = case_when(is.na(group) == TRUE ~ "non_significant", TRUE ~ group))

plot <- ggplot(plot_df, aes(x = generation,
             y = variance_log2CPM)) +
  geom_boxplot() +
  theme_classic() + 
    xlab("Generation") +
  scale_x_discrete(labels = c("0" ,"7", "31")) +
theme(
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="black",
                                    fill="white")) +
  coord_cartesian(ylim = c(0,0.05)) +
  facet_wrap(~ group)

## Compare means across groups - exclude non-significant genes
plot_df <- plot_df %>% dplyr::filter(group != "non_significant") %>% droplevels()
ggpubr::compare_means(variance_log2CPM ~ generation, plot_df, group.by = "group")


```
## Boxplots - all DEGs non-Plateau genes
```{r}
plot_df <- variance_per_gene_df %>%
  mutate(group = case_when(is.na(group) == TRUE ~ "non_significant", TRUE ~ group))
plot_df <- plot_df %>% dplyr::filter(!group %in% c("non_significant", "plateau")) %>% droplevels()
 
# Relevel group:
plot_df$group <- factor(plot_df$group, levels = c("complete_reversed",
                                                  "incomplete_reversed",
                                                  "late_response",
                                                  "monotonic"))

plot_df <- plot_df %>% mutate(group2 = case_when(group %in% c("complete_reversed", "incomplete_reversed") ~ "A",
                                                 group %in% c("late_response", "monotonic") ~ "B"))

plot_df %>% ggplot(aes(y = variance_log2CPM, x = generation, color = group)) +
  geom_boxplot(linewidth=0.6) +  
  scale_color_manual(values = c(color_complete_reversed,
                                color_incomplete_reversed,
                                color_late_response,
                                color_monotonic),
                     labels = c("Complete-reverse",
                                "Incomplete-reverse",
                                "Late-response",
                                "Monotonic")) +  
  theme_classic() +
  ylab(bquote(Gene~expression~(log[2]*CPM)~variance)) + 
  xlab("Generation") +
  theme(legend.position = "bottom")  +
  coord_cartesian(ylim = c(0, 0.3)) +
  scale_x_discrete(labels = c("0" ,"7", "31")) +
theme(
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="black",fill="white"),
    axis.text.x.top  = element_blank(),
    axis.ticks.x.top = element_blank(),
    axis.line.x.top = element_blank()) +
   facet_wrap(~ group2 + direction, ncol=4, strip.position="top") + scale_y_break(c(0.05,0.27), scales = c(0.2,2)) 

# Export plot:
#ggsave("output_files/plots/correlation_plots_and_trajectories/gene_variance_groups.svg",
       #width = 300,
       #height = 200,
       #units = "mm",
       #dpi = 300, last_plot())

```
# Wilkoxon test distribution of gene variances (up and down) - 4 groups 
Here we compare the variance of some groups within generation.
I perform different tests for up-regulated and down-regulated genes.
We are only interested in:
complete-reverse vs incomplete-reverse, late-response and monotonic.
```{r}

# generation:
generation <- c("gen07", "gen31")
# logFC direction:
direction <- c("up", "down")

# data_filter_df: keep needed groups
data_filter_A_df <- variance_per_gene_df %>%
  dplyr::filter(group %in% c("complete_reversed",
                             "incomplete_reversed")) 

# Create an empty data frame to store results
wilcoxon_results_A <- data.frame(
  generation = character(),
  #direction = character(),
  group1 = character(),
  group2 = character(),
  p_value = numeric(),
  stringsAsFactors = FALSE)

for (gen in unique(data_filter_A_df$generation)) {
  for (direc in unique(data_filter_A_df$direction)) {
    
    # Subset the data for the current generation and direction
    subset_data <- data_filter_A_df %>%
      filter(generation == gen, direction == direc)
    
    # Get the unique full_groups for pairwise comparison
    full_groups <- unique(subset_data$full_group)
    
    # Loop over all pairs of full_groups
    for (i in 1:(length(full_groups) - 1)) {
      for (j in (i + 1):length(full_groups)) {
        
        group1 <- full_groups[i] 
        group2 <- full_groups[j] 
        
        # Subset the data for the two groups
        group1_data <- subset_data[subset_data$full_group == group1, "variance_log2CPM"] 
        group2_data <- subset_data[subset_data$full_group == group2, "variance_log2CPM"]
        
        # Perform Wilcoxon test
        test_result <- wilcox.test(group1_data, group2_data)
        
        # Store the results in the data frame
        wilcoxon_results_A <- rbind(wilcoxon_results_A,
          data.frame(test = "Wilcoxson rank sum test, two tailed",
          generation = gen,
          logFC_direction = direc,
          group1 = group1,
          group2 = group2,
          p_value = test_result$p.value))
      }
    }
  }
}

###

# Create an empty data frame to store results
wilcoxon_results_B <- data.frame(
  generation = character(),
  #direction = character(),
  group1 = character(),
  group2 = character(),
  p_value = numeric(),
  stringsAsFactors = FALSE)


data_filter_B_df <- variance_per_gene_df %>%
  dplyr::filter(group %in% c("monotonic", "late_response")) %>% droplevels()

for (gen in unique(data_filter_B_df$generation)) {
  for (direc in unique(data_filter_B_df$direction)) {
    
    # Subset the data for the current generation and direction
    subset_data <- data_filter_B_df %>%
      filter(generation == gen, direction == direc)
    
    # Get the unique full_groups for pairwise comparison
    full_groups <- unique(subset_data$full_group)
    
    # Loop over all pairs of full_groups
    for (i in 1:(length(full_groups) - 1)) {
      for (j in (i + 1):length(full_groups)) {
        
        group1 <- full_groups[i] 
        group2 <- full_groups[j] 
        
        # Subset the data for the two groups
        group1_data <- subset_data[subset_data$full_group == group1, "variance_log2CPM"] 
        group2_data <- subset_data[subset_data$full_group == group2, "variance_log2CPM"]
        
        # Perform Wilcoxon test
        test_result <- wilcox.test(group1_data, group2_data)
        
        # Store the results in the data frame
        wilcoxon_results_B <- rbind(wilcoxon_results_B,
          data.frame(test = "Wilcoxson rank sum test, two tailed",
          generation = gen,
          logFC_direction = direc,
          group1 = group1,
          group2 = group2,
          p_value = test_result$p.value))
      }
    }
  }
}

# Combine results:
wilcoxon_results <- rbind(wilcoxon_results_A, wilcoxon_results_B)

# Adjust p-values across all tests:
wilcoxon_results <- wilcoxon_results %>%
  #group_by(generation, logFC_direction) %>% 
  mutate(padj_fdr = p.adjust(p_value, method = "fdr"))

# Include asterisks:
wilcoxon_results <- wilcoxon_results %>%
  mutate(significance = case_when(
    padj_fdr >= 0.05 ~ "n.s",          
    padj_fdr <= 0.05 & padj_fdr > 0.01 ~ "*",  
    padj_fdr <= 0.01 & padj_fdr >= 0.001 ~ "**", 
    padj_fdr < 0.001 ~ "***",
    padj_fdr <= 0.0001 ~ "****"))

# Export results:
export_df <- wilcoxon_results
#write_xlsx(export_df, "/output_files/S3_gene_variance_across_groups_tests.xlsx")

```

--- END ---