---
title: "High-Protein BigCages - Transcriptomic parallelism"
Author: "Claudia Ramirez-lanzas"
Date: "27/05/2024"
---

# Load libraries:
```{r}
# Data handling:
library("writexl") 
library("tidyverse") 

# Plotting:
library("ggplot2") 
library("svglite") 
library("UpSetR") # UpSet plots
library("patchwork") 
library("khroma") # plotting colour blind safe
library("ggpubr") 
library("rstatix")
library("ggsignif")



```

# Check sessionInfo:
```{r}
sessionInfo()
```
# Set-up colors:
```{r}
bright <- color("bright")
plot_scheme(bright(6), colours = TRUE, names = TRUE, size = 0.9)
#
muted <- color("muted")
plot_scheme(muted(9), colours = TRUE, names = TRUE, size = 0.9)

color_plateau <- muted(9)[2] #"#332288"
color_monotonic <- muted(9)[7] #"#44AA99"
color_late_response <-  muted(9)[3] #"#DDCC77"
color_incomplete_reversed <- muted(9)[9] #"#AA4499"
color_complete_reversed <- muted(9)[5] #"#88CCEE" 
color_shared <- muted(9)[1] #"#CC6677" 

plot_scheme(muted(9)[2], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[7], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[9], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[3], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[5], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[1], colours = TRUE, names = TRUE, size = 0.9)
```
# Import files:
```{r}
DE_rep_specific_results <- read_csv("input_files/DE_repwise_results.csv", show_col_types = FALSE)
DE_rep_specific_results <- DE_rep_specific_results %>% dplyr::filter(generation != "transitory_F7_F31") 

DE_overall_results <- read_csv("input_files/DE_overall_results.csv", show_col_types = FALSE)

```

*DE_rep_specific_results: contains the results of replicate wise DE analysis. 
P-values are corrected across all contrasts.


* DE_overall_results: contains the results of DE analysis using dream model. 
Including 3 contrasts: Base vs F7, Base vs F31 and F7 vs F31. P-values corrected across the 3 contrasts using the fdr method.
# Function 1: Compute Jaccard Index per group
This function uses replicate wise DE results to compute Jaccard Index in >= 1 groups.
Arguments:

# Function 1: Compute Jaccard Index
* df: a data-set containing at least 3 columns:
1) the gene_id of the DEGs (only include DEGs, non-significant genes should be
excluded here), 2)
the replicate id, 3)
the group id (e.g., the transcriptomic group, the generation etc)

* group: the column name for the grouping variable 
```{r}
compute_jaccard_index_per_group <- function(df, group) {
  # Summarize gene IDs per contrast
  gene_ids_per_contrast <- df %>%
    group_by(group, contrast) %>%
    summarise(gene_ids = list(gene_id), .groups = 'drop')
  
  # Create an empty data-frame to store the results:
  results <- data.frame(
    group = integer(),
    contrast1 = character(),
    contrast2 = character(),
    Jaccard_Index = numeric(),
    stringsAsFactors = FALSE)
  
  # Generate pairwise comparisons:
  unique_contrasts <- unique(gene_ids_per_contrast$contrast)
  pairwise_comparisons <- combn(unique_contrasts, 2, simplify = FALSE)
  
  # Create a data frame to store these comparisons:
  comparison_list <- data.frame(
    contrast1 = character(),
    contrast2 = character(),
    stringsAsFactors = FALSE)
  
  # Introduce pair wise comparisons:
  for (pair in pairwise_comparisons) {
    comparison_list <- rbind(comparison_list, data.frame(
      contrast1 = pair[1],
      contrast2 = pair[2],
      stringsAsFactors = FALSE
    ))
  }
  
  # Perform comparisons based on the comparison list:
  for (grp in unique(gene_ids_per_contrast$group)) {
    # Subset data for the current group:
    subset_data <- gene_ids_per_contrast %>% filter(group == grp)
    
    for (i in 1:nrow(comparison_list)) {
      contrast1 <- comparison_list$contrast1[i]
      contrast2 <- comparison_list$contrast2[i]
      
      # Get gene IDs for each contrast:
      gene_ids1 <- subset_data %>% filter(contrast == contrast1) %>% pull(gene_ids) %>% unlist()
      gene_ids2 <- subset_data %>% filter(contrast == contrast2) %>% pull(gene_ids) %>% unlist()
      
      # Calculate Jaccard index:
      jaccard_index <- length(intersect(gene_ids1, gene_ids2)) / length(union(gene_ids1, gene_ids2))
      
      # Store the result:
      results <- rbind(results, data.frame(
        group = grp,
        contrast1 = contrast1,
        contrast2 = contrast2,
        Jaccard_Index = jaccard_index,
        stringsAsFactors = FALSE
      ))
    }
  }
  
  return(results)
}

```
# Function 2: Compute pairwise Pearson´s correlation:
This function uses replicate wise DE results to compute Pearson correlation on
the whole transcriptome log2FC responses. 

*data: the replicate wise DE results containing the log2FC of all replicates. 
*generation: generation that we want to use 
*logFC_generation: column name containing generation specific log2FC
```{r}
compute_correlations <- function(data, generation, logFC_generation) {
  
  data <- filtered_df %>%  dplyr::filter(generation == !!generation) %>%
    select(gene_id, !!logFC_generation, replicate) 
  # Pivot data so that replicates are columns and logFC values are rows
  wide_data <- tidyr::pivot_wider(data, names_from = replicate, values_from = logFC_generation)
  
  # Remove the generation column for correlation computation
  wide_data <- wide_data %>% dplyr::select(gene_id, R1, R2, R3, R4, R5, R6)
  
  # Compute Pearson correlation for pairwise replicates
  cor_matrix <- stats::cor(wide_data[, c("R1", "R2", "R3", "R4", "R5", "R6")])
  
  # Create a data_frame:
  cor_df <- as.data.frame(as.table(cor_matrix))
  cor_df$contrast <- paste(cor_df$Var1, cor_df$Var2, sep = "_")
  cor_df$generation <- generation
  cor_df$analysis <- "Pearson´s correlation on log2FC"
  cor_df <- cor_df %>% dplyr::select(generation, analysis, contrast, Freq)
  colnames(cor_df) <- c("generation", "analysis", "contrast", "pearson_corr")
  
  # Round pearson correlation coefficient up to 3 decimals:
  cor_df$pearson_corr <- round(cor_df$pearson_corr, 3)
  
  return(cor_df)
}

```

```{r}
compute_correlations <- function(data, generation) {
  
  data <- filtered_df %>%  dplyr::filter(generation == !!generation) %>%
    select(gene_id, logFC, replicate) 
  # Pivot data so that replicates are columns and logFC values are rows
  wide_data <- tidyr::pivot_wider(data, names_from = replicate, values_from = logFC)
  
  # Remove the generation column for correlation computation
  wide_data <- wide_data %>% dplyr::select(gene_id, R1, R2, R3, R4, R5, R6)
  
  # Compute Pearson correlation for pairwise replicates
  cor_matrix <- stats::cor(wide_data[, c("R1", "R2", "R3", "R4", "R5", "R6")])
  
  # Create a data_frame:
  cor_df <- as.data.frame(as.table(cor_matrix))
  cor_df$contrast <- paste(cor_df$Var1, cor_df$Var2, sep = "_")
  cor_df$generation <- generation
  cor_df$analysis <- "Pearson´s correlation on log2FC"
  cor_df <- cor_df %>% dplyr::select(generation, analysis, contrast, Freq)
  colnames(cor_df) <- c("generation", "analysis", "contrast", "pearson_corr")
  
  # Round pearson correlation coefficient up to 3 decimals:
  cor_df$pearson_corr <- round(cor_df$pearson_corr, 3)
  
  return(cor_df)
}

```
# # Part 1: Replicate especific results:
## Summary plots:
```{r}

DE_rep_specific_results %>% dplyr::filter(padj_fdr < 0.05) %>% 
  ggplot(aes(x = contrast, fill = generation)) + 
  geom_bar() + 
  scale_fill_manual(values = c("#af8dc3","#762a83")) +
  geom_text(stat='count', aes(label= after_stat(count)), position = "stack") + 
  labs(x = "replicate", y = "N° DEGs", title = "") + 
  theme(axis.title = element_text(size = 15))  + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(labels = c("R1", "R2", "R3", "R4", "R5", "R6", 
                              "R1", "R2", "R3", "R4", "R5", "R6")) + 
  facet_wrap(~ generation,
             scales = "free_x",
             labeller = labeller(generation = c("gen07" = "F7", 
                                                "gen31" = "F31"))) + 
  guides(fill = "none") +
  theme_classic() 


```
## UpSet plots:
```{r}
#Create a list with DEGs of each replicate:
list_upset_plot <- list(
  rep_1_F7 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen07" & padj_fdr < 0.05 & contrast == "L_base_01_F7") %>%
    dplyr::pull(gene_id), 
  rep_2_F7 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen07" & padj_fdr < 0.05 & contrast == "L_base_02_F7") %>%
    dplyr::pull(gene_id),
  rep_3_F7 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen07" & padj_fdr < 0.05 & contrast == "L_base_03_F7") %>%
    dplyr::pull(gene_id),
  rep_4_F7 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen07" & padj_fdr < 0.05 & contrast == "L_base_04_F7") %>%
    dplyr::pull(gene_id),
  rep_5_F7 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen07" & padj_fdr < 0.05 & contrast == "L_base_05_F7") %>%
    dplyr::pull(gene_id),
  rep_6_F7 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen07" & padj_fdr < 0.05 & contrast == "L_base_06_F7") %>%
    dplyr::pull(gene_id),
  rep_1_F31 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen31" & padj_fdr < 0.05 & contrast == "L_base_01_F31") %>%
    dplyr::pull(gene_id), 
  rep_2_F31 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen31" & padj_fdr < 0.05 & contrast == "L_base_02_F31") %>%
    dplyr::pull(gene_id),
  rep_3_F31 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen31" & padj_fdr < 0.05 & contrast == "L_base_03_F31") %>%
    dplyr::pull(gene_id),
  rep_4_F31 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen31" & padj_fdr < 0.05 & contrast == "L_base_04_F31") %>%
    dplyr::pull(gene_id),
  rep_5_F31 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen31" & padj_fdr < 0.05 & contrast == "L_base_05_F31") %>%
    dplyr::pull(gene_id),
  rep_6_F31 = DE_rep_specific_results %>%
    dplyr::filter(generation == "gen31" & padj_fdr < 0.05 & contrast == "L_base_06_F31") %>%
    dplyr::pull(gene_id))

# Combine all gene IDs into a single list:
list_upset_plot2 <- sapply(list_upset_plot, identity)
a <- list_upset_plot2[1:6] # F7
b <- list_upset_plot2[7:12] # F31
c <- list_upset_plot2 # F7 and F31

names(a) <- c("rep 1-gen 7",
                    "rep 2-gen 7",
                    "rep 3-gen 7",
                    "rep 4-gen 7",
                    "rep 5-gen 7",
                    "rep 6-gen 7")

names(b) <- c("rep 1-gen 31",
                    "rep 2-gen 31",
                    "rep 3-gen 31",
                    "rep 4-gen 31",
                    "rep 5-gen 31",
                    "rep 6-gen 31")

names(c) <- c("rep 1-gen 7",
                    "rep 2-gen 7",
                    "rep 3-gen 7",
                    "rep 4-gen 7",
                    "rep 5-gen 7",
                    "rep 6-gen 7",
                    "rep 1-gen 31",
                    "rep 2-gen 31",
                    "rep 3-gen 31",
                    "rep 4-gen 31",
                    "rep 5-gen 31",
                    "rep 6-gen 31")

#upset_plot_F7: 

# Data-set containing all unique DEGs:
df1 <- data.frame(gene=unique(unlist(a)))
# Data-set containing the genes associated to each set:
df2 <-  lapply(a, function(x){
  data.frame(gene = x)
}) %>% 
  bind_rows(.id = "path") 
# Data-set containing the genes associated to each set:
df_int_F7 <- lapply(df1$gene,function(x){
  # pull the name of the intersections
  intersection <- df2 %>% 
    dplyr::filter(gene==x) %>% 
    arrange(path) %>% 
    pull("path") %>% 
    paste0(collapse = "|")
  # build the data-frame
  data.frame(gene_id = x,
             intersection = intersection)
}) %>% 
  bind_rows()
df_int_F7$intersection <- gsub(".gene_id", "", df_int_F7$intersection )

# Summarize the counts per intersection:
summ_F7 <- df_int_F7 %>% 
  dplyr::group_by(intersection) %>% 
  dplyr::summarise(n=n()) %>% 
  dplyr::arrange(desc(n))
# Plot upset plot:
upset_plot_F7 <- UpSetR::upset(fromList(a),
                                nsets = length(a),
                                nintersects = 200,
                                order.by = "degree",
                                text.scale = c(2, 2, 2, 2, 2, 1.3))  

#upset_plot_F31: 

#upset_plot_F31: 
# Data-set containing all unique DEGs:
df1 <- data.frame(gene=unique(unlist(b)))
# Data-set containing the genes associated to each set:
df2 <-  lapply(b, function(x){
  data.frame(gene = x)
}) %>% 
  bind_rows(.id = "path") 
# Data-set containing the genes associated to each set:
df_int_F31 <- lapply(df1$gene,function(x){
  # pull the name of the intersections
  intersection <- df2 %>% 
    dplyr::filter(gene==x) %>% 
    arrange(path) %>% 
    pull("path") %>% 
    paste0(collapse = "|")
  # build the data-frame
  data.frame(gene_id = x, intersection = intersection)
}) %>% 
  bind_rows()
df_int_F31$intersection <- gsub(".gene_id", "", df_int_F31$intersection )

# Summarize the counts per intersection:
summ_F31 <- df_int_F31 %>% 
  dplyr::group_by(intersection) %>% 
  dplyr::summarise(n=n()) %>% 
  dplyr::arrange(desc(n))

# Plot upset plot:
upset_plot_F31 <- UpSetR::upset(fromList(b),
                                nsets = length(b),
                                nintersects = 200,
                                order.by = "degree",
                                text.scale = c(2, 2, 2, 2, 2, 1.3)) 

upset_plot_F7
upset_plot_F31


```
text.scale = c(1.3, 1.3, 1.3, 1.3, 1.3, 0.75)
intersection size title,
intersection size tick labels,
set size title,
set size tick labels,
set names,
and numbers above bars.

## Aggregated Up-set plot
```{r}
## F7
# Create a new column 'rep_count'
summ_F7_count <- summ_F7 %>%
  mutate(rep_count = str_count(intersection, "rep"))
# Summarize the counts per intersection2:
summ_F7_2 <- summ_F7_count %>%
  group_by(rep_count) %>%
  summarize(total_n = sum(n))

## F31
# Create a new column 'intersection2'
summ_F31_count <-  summ_F31 %>%
  mutate(rep_count = str_count(intersection, "rep"))
# Summarize the counts per intersection2:
summ_F31_2 <- summ_F31_count %>%
  group_by(rep_count) %>%
  summarize(total_n = sum(n))


# Add generation:
summ_F7_2$generation <- "7"
summ_F31_2$generation <- "31"

# Combine generations:
summ_F7_F31 <- rbind(summ_F7_2, summ_F31_2)

# Relabel rep_count:
summ_F7_F31 <- summ_F7_F31 %>% dplyr::mutate(sharing = case_when(rep_count == 1 ~ "private",
                                                       rep_count == 2 ~ "shared_by_2",
                                                       rep_count == 3 ~ "shared_by_3",
                                                       rep_count == 4 ~ "shared_by_4",
                                                       rep_count == 5 ~ "shared_by_5",
                                                       rep_count == 6 ~ "shared_by_6"))
# Relevel sharing:
summ_F7_F31$sharing <- factor(summ_F7_F31$sharing,
                                  levels = c("shared_by_6", "shared_by_5",
                                             "shared_by_4", "shared_by_3",
                                             "shared_by_2", "private"))

# Relevel generation:
summ_F7_F31$generation <- factor(summ_F7_F31$generation,
                                  levels = c("7", "31"))
# Plotting:
aggregated_upset_plot <- summ_F7_F31 %>%
  ggplot(aes(x = sharing, y= total_n, color = generation)) +
  geom_bar(stat = "identity",
           fill = "white",
           linewidth =2) +
  geom_text(aes(label = total_n), vjust = -0.5, size = 3, color = "black") +  
  labs(title = "",
       x = "No of replicates with expression changes",
       y = "N. DEGs") +
  scale_color_manual(values = c("#af8dc3","#762a83")) + 
  scale_x_discrete(labels=c("shared_by_6" = "6",
                            "shared_by_5" = "5",
                            "shared_by_4" = "4",
                            "shared_by_3" = "3",
                            "shared_by_2" = "2",
                            "private" = "1")) +
  theme_linedraw() +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="white",
                                    fill="white")) +
  facet_wrap(~generation)

aggregated_upset_plot

# Save the plot:
#ggsave(plot = aggregated_upset_plot,
       #"output_files/plots/aggregated_upset_plots.svg",
       #height = 180,
       #width = 180, 
       #units = "mm", 
       #dpi = 300)

```
# Create shared_upset_data:
```{r}
# Merge F7 and F31:

a <- df_int_F7
a$generation <- "gen07"
b <- df_int_F31
b$generation <- "gen31"

shared_upset_data <- rbind(a, b)

# Prepare the data-set:

shared_upset_data <- shared_upset_data %>%
  mutate(rep_count = str_count(intersection, "rep"))

shared_upset_data <- shared_upset_data %>% dplyr::mutate(sharing = case_when(rep_count == 1 ~ "private",
                                                       rep_count == 2 ~ "shared_by_2",
                                                       rep_count == 3 ~ "shared_by_3",
                                                       rep_count == 4 ~ "shared_by_4",
                                                       rep_count == 5 ~ "shared_by_5",
                                                       rep_count == 6 ~ "shared_by_6"))
```
# Export shared_upset_data:
```{r}
export_df <- shared_upset_data
#write_csv(export_df, "output_files/shared_upset_data.csv")
#write_csv(export_df, "input_files/shared_upset_data.csv")

```

# Create DE_rep_specific_results_filt:
```{r}
# Remove non-significant genes from replicate-specific DE data-set:
DE_rep_specific_results_filt <- DE_rep_specific_results %>% dplyr::filter(padj_fdr < 0.05)

```
# # Part 2: Parallelism - Jaccard Index:
## Subset data for generation 7 and 31:
```{r}
df_F7 <- DE_rep_specific_results_filt %>% dplyr::filter(generation == "gen07") %>%
  dplyr::select(c(gene_id, contrast, generation))
# Rename generation as group to use the function in next step:
colnames(df_F7) <- c("gene_id", "contrast", "group")

df_F31 <- DE_rep_specific_results_filt %>% dplyr::filter(generation == "gen31") %>%
  dplyr::select(c(gene_id, contrast, generation))
# Rename generation as group to use the function in next step:
colnames(df_F31) <- c("gene_id", "contrast", "group")

```

## ##Jaccard Index F7 vs F31: 
## Compute Jaccard Index per generation:
```{r}
# Here the group is the generation so:
df_F7_Jaccard <- compute_jaccard_index_per_group(df_F7, group) 
df_F7_Jaccard$gen <- "gen07"
 
df_F31_Jaccard <- compute_jaccard_index_per_group(df_F31, generation)
df_F31_Jaccard$gen <- "gen31"

Jacccard_F7_F31_results <- rbind(df_F7_Jaccard, df_F31_Jaccard)


```
## Mean Jaccard Index:
```{r}
Jacccard_F7_F31_results %>%
  group_by(gen) %>%
  summarise_at(vars(Jaccard_Index), funs(mean(., na.rm=TRUE)))

```
## Fig. 5 B - Plot Jaccard Index:
```{r}
plot_df <- Jacccard_F7_F31_results


plot_df$Jaccard_Index <- as.numeric(plot_df$Jaccard_Index)
plot_df$generation <- factor(plot_df$gen, levels = c("gen07","gen31"))

plot_JI_F7_F31 <- plot_df %>% 
  ggplot(aes(x = gen, y = round(Jaccard_Index, digits = 4))) + 
  geom_boxplot(aes(color = generation, width=0.3),
               show.legend = FALSE) +
  geom_point(aes(color = generation), size = 2, show.legend = FALSE) + 
  labs(x = "Generation", y = "Jaccard index", title = "") + 
    scale_x_discrete(labels = c("7", "31")) +
  theme_linedraw() +
  scale_color_manual(values = c("#af8dc3", "#762a83")) + 
theme( aspect.ratio = 1,
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 16, color = "black"),
    axis.text.y = element_text(size = 16, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="white",
                                    fill="white")) +
  ylim(0, 1)  


## Add statistics to the plot ##
plot_df$Jaccard_Index <- as.numeric(plot_df$Jaccard_Index)
plot_df$generation <- as.factor(plot_df$generation)

plot_JI_F7_F31 <- plot_JI_F7_F31 +
  stat_compare_means(method = "wilcox.test",
                     label = "p.format", hjust = -2)

plot_JI_F7_F31


# Save the plot:
#ggsave(plot = last_plot(),
       #"output_files/plots/Jaccard_Index_and_parallelism/Jaccard_Index_F7_vs_F31.svg", 
       #width = 180,
       #height = 100,
       #units = "mm", 
       #dpi = 300)



```

# ## Part 3: Pearson´s correlation
## Prepare the dataset:
```{r}

DE_rep_specific_results %>%
  group_by(generation, contrast) %>%
  summarise(count = n())
# Different sized transcriptomes are shown across replicates because their
#datasets were filtered by low expression independently.
# So, I will create an intersection for them:
# Find the intersection of gene_ids across all replicates:
common_genes <- DE_rep_specific_results %>%
  group_by(gene_id) %>%
  summarise(rep_count = n_distinct(contrast)) %>%
  filter(rep_count == 12) %>%
  pull(gene_id) # 10,197 genes
# Filter data-set to keep the intersected genes
#(i.e., transcriptome after filtering by low expression in all replicates):
filtered_df <- DE_rep_specific_results %>% filter(gene_id %in% common_genes)
filtered_df %>%
  group_by(generation, contrast) %>%
  summarise(count = n())

# Include column replicate:

filtered_df <- filtered_df %>% mutate(replicate =  case_when(
  grepl("_01_", contrast) ~ "R1",
  grepl("_02_", contrast) ~ "R2",
  grepl("_03_", contrast) ~ "R3",
  grepl("_04_", contrast) ~ "R4",
  grepl("_05_", contrast) ~ "R5",
  grepl("_06_", contrast) ~ "R6"))

```
## Compute pairwise correlations within generation
```{r}
pearson_corr_F7 <- compute_correlations(filtered_df, generation = "gen07")
pearson_corr_F31 <- compute_correlations(filtered_df, generation = "gen31")

pearson_corr_df <- rbind(pearson_corr_F7, pearson_corr_F31)
```

## Fig 5 C - plor pearson correlation
```{r}
# Exclude correlations between same replicate:
plot_df <- pearson_corr_df %>% dplyr::filter(!contrast %in% c("R1_R1",
                                                              "R2_R2",
                                                              "R3_R3",
                                                              "R4_R4",
                                                              "R5_R5",
                                                              "R6_R6")) 

plot_df %>% ggplot(aes(x = pearson_corr, y =generation, fill = generation)) +
  geom_violin(alpha=1) +
  geom_boxplot(fill = "white", width=0.15, color="black", alpha=0.5, outlier.size = 0) +
  scale_fill_manual(values = c("#af8dc3", "#762a83")) +
  theme_linedraw() +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 16, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.text.x.top = element_text(colour="black", face = "bold", size = 16),
    strip.background = element_rect(colour="white",
                                    fill="white")) + 
  xlab("Pearson's correlation coefficient") 

# Save the plot:
#ggsave(plot = last_plot(),
       #"output_files/plots/Jaccard_Index_and_parallelism/pearson_correlation.svg", 
       #width = 200,
       #height = 150,
       #units = "mm", 
       #dpi = 400)


```
# ##Part 4: distribution of parallelism in overall-model
# Classify all overall DEGs according to rep-wise categories:
```{r}
# DEGs found by the overall model:
F7 <- DE_overall_results %>% dplyr::filter(generation =="gen07" & padj_fdr < 0.05)
# DEGs found by the rep-specific models:
F7_rs <- shared_upset_data %>% dplyr::filter(generation == "gen07")
# DEGs found by the rep-specific and overall models:
F7_rs_ov <- dplyr::left_join(F7, F7_rs, by = c("generation", "gene_id"))
# Exclude some genes that were filtered by low expression in the overall-model 
F7_rs_ov <- na.omit(F7_rs_ov)

##
# DEGs found by the overall model:
F31 <- DE_overall_results %>% dplyr::filter(generation =="gen31" & padj_fdr < 0.05)
# DEGs found by the rep-specific models:
F31_rs <- shared_upset_data %>% dplyr::filter(generation == "gen31")
# DEGs found by the rep-specific and overall models:
F31_rs_ov <- dplyr::left_join(F31, F31_rs, by = c("generation", "gene_id"))
# Exclude some genes that were filtered by low expression in the overall-model:
F31_rs_ov <- na.omit(F31_rs_ov)

# Combine generations:
summ_F7_F31 <- rbind(F7_rs_ov, F31_rs_ov)
# Relevel sharing:
summ_F7_F31$sharing <- factor(summ_F7_F31$sharing,
                                  levels = c("shared_by_6", "shared_by_5",
                                             "shared_by_4", "shared_by_3",
                                             "shared_by_2", "private"))
# Relevel generation:
summ_F7_F31$generation <- factor(summ_F7_F31$generation,
                                  levels = c("gen07", "gen31"))

summ_F7_F31 %>%
  ggplot(aes(x = sharing, color = generation)) +
  geom_bar(fill = "white",
           linewidth =2) +
geom_text(stat = 'count', aes(label = ..count..),
            vjust = -0.5, size = 5, color = "black") +
  labs(title = "",
       x = "No of replicates with expression changes",
       y = "No DEGs") +
  scale_color_manual(values = c("#af8dc3","#762a83")) +
  scale_x_discrete(labels = c("6","5","4","3","2","1")) +
  theme_linedraw() +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16, color = "black"),
    axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position="none",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.background=element_rect(colour="white",
                                    fill="white")) +
  facet_wrap(~generation) 

# Save plot:
#ggsave("output_files/plots/Jaccard_Index_and_parallelism/rep_freq_spectrum_overall_vs_repwise.svg", last_plot())

```

--- END ---