---
title: "High-Protein BigCages - relationship of pleiotropy and parallelism"
Author: "Claudia Ramirez-lanzas"
Date: "16/01/2025"
---

# Load libraries:
```{r}
# Data handling:
library("readxl") 
library("writexl") 
library("tidyverse")

# Plotting:
library("ggplot2") 
library("svglite") # plotting
library("patchwork") 
library("gridExtra")
library("ggpubr") # add p-values to plots
library("khroma") # plotting colour blind safe

```
# Check sessionInfo:
```{r}
sessionInfo()
```
# Set-up colors:
```{r}
bright <- color("bright")
plot_scheme(bright(6), colours = TRUE, names = TRUE, size = 0.9)
#
muted <- color("muted")
plot_scheme(muted(9), colours = TRUE, names = TRUE, size = 0.9)

color_plateau <- muted(9)[2] #"#332288"
color_monotonic <- muted(9)[7] #"#44AA99"
#color_late_response <-  muted(9)[3] #"#DDCC77"
color_late_response <-  "#D4AF37"
color_incomplete_reversed <- muted(9)[9] #"#AA4499"
color_complete_reversed <- muted(9)[5] #"#88CCEE" 
color_shared <- muted(9)[1] #"#CC6677" 

plot_scheme(muted(9)[2], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[7], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[9], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[3], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[5], colours = TRUE, names = TRUE, size = 0.9)
plot_scheme(muted(9)[1], colours = TRUE, names = TRUE, size = 0.9)


```
# Import files:
```{r}
connectivity_df <- read_csv("input_files/connectivity_per_gene.csv",  show_col_types = FALSE)

tau_df <- read_csv("input_files/oneminusTau_per_gene.csv",  show_col_types = FALSE)

shared_upset_data <- read_csv("input_files/shared_upset_data.csv", show_col_types = FALSE)

DE_rep_wise_results <- read_csv("input_files/DE_repwise_results.csv", show_col_types = FALSE)

```

* connectivity_df : contains gene-wise information about the log2FC at each
generation, the transcriptomic group and also the connectivity value. This dataset 
was generated in 12_pleiotropy

* tau_df : contains gene-wise information about the log2FC at each generation,
the transcriptomic group and also the 1-Tau (called tau) value. This dataset 
was generated in 12_pleiotropy

* shared_upset_data: contains the data-set used to make the UpSet plots.
It contains information about the DEGs detected in the replicate specific DE
analysis. That is, each replicate was compared to the Base and the DEGs were
compared to the other 5 replicates leading to 6 categories of parallelism
(ranging from 1 (private DEGs) to 6 (shared by all replicates)).

* DE_rep_wise_results: contains the results of DE analysis using dream model. 
6 independent modeles, one per replicate. Including 3 generations.
Each replicated had 3 samples. P-values corrected across all contrasts and replicates. 

# Create parallel_pleio_df:
```{r}

# Subset Background genes (all genes with connectivity information):
background_conn <- connectivity_df %>% dplyr::filter(group == "background")
# Include connectivity to the parallelism data:
parallel_pleio_df <- full_join(shared_upset_data, background_conn, by = "gene_id")

# Include tau to the parallelism data:

# Subset Background genes (all genes with connectivity information):
background_tau <- tau_df %>% dplyr::filter(group == "background")
background_tau <- background_tau %>% dplyr::select(gene_id, tau)
parallel_pleio_df <- left_join(parallel_pleio_df, background_tau, by = "gene_id")


# Keep only genes with concordant changes across all replicates:
# Label Concordant and Non_concordant_genes:
DE_rep_specific_DEGs <- DE_rep_wise_results %>%
  mutate(generation = case_when(grepl("F7", contrast) ~ "gen7",
                                grepl("F31", contrast) ~ "gen31")) %>%
  dplyr::filter(padj_fdr < 0.05)
concordant_DEGs <- DE_rep_specific_DEGs %>% group_by(generation,gene_id) %>%
  mutate(
    concordance = ifelse(all(sign(logFC) == 1) | all(sign(logFC) == -1),
                         "concordant", "non-concordant")) %>% ungroup()
concordant_DEGs <- concordant_DEGs %>% dplyr::filter(concordance == "concordant")
table(concordant_DEGs$concordance)

# Keep only concordant genes:
parallel_pleio_df <- parallel_pleio_df %>% dplyr::filter(gene_id %in% concordant_DEGs$gene_id)

# Remove private genes:
parallel_pleio_df <- parallel_pleio_df %>% dplyr::filter(rep_count != 1)

```


# Part 1: Correlation between pleiotropy and parallellism - using only replicate especific DE dataset
Here we correlate gene-wise pleiotropic levels with gene-wise parallelism levels.
Here we exclude private DEGs -> those DEGs found in 1 replicate
Here we exclude non-concordant DEGs <- those with different signs (log2FC) across at least 1 replicate.

# Plot 1
```{r}
parallel_pleio_df %>% ggplot(aes(y=log10_connectivity, x=factor(rep_count), fill=factor(rep_count))) +
  geom_boxplot() +
  scale_fill_manual(values =c("#fdd49e","#fc8d59","#ef6548","#d7301f","#7f0000")) +
  ylab(bquote(log[10]*(connectivity))) +
  xlab("Parallelism (No of replicates with expression changes)") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.ticks = element_line(color="black"),
    axis.text = element_text(size = 14, color = "black"),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position ="none",
    panel.border = element_rect(colour = "black",
                                fill=NA,
                                linewidth=0.5)) +
  facet_wrap(~generation)



##


connec_parallel_plot <- parallel_pleio_df %>% ggplot(aes(y=log10_connectivity, x=factor(rep_count), fill=factor(rep_count))) +
  geom_violin() +
  geom_boxplot(fill = "white", width=0.2, color="black", alpha=0.4, outlier.size = 0) +
  scale_fill_manual(values =c("#fdd49e","#fc8d59","#ef6548","#d7301f","#7f0000")) +
  ylab(bquote(log[10]*(connectivity))) +
  xlab("Parallelism (No of replicates with expression changes)") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.ticks = element_line(color="black"),
    axis.text = element_text(size = 14, color = "black"),
    strip.text.x.top = element_text(colour="black", face = "bold", size = 16),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position ="none",
    panel.border = element_rect(colour = "black",
                                fill=NA,
                                linewidth=0.5),
    strip.background = element_rect(colour="white",fill="white")) +
  facet_wrap(~ generation, labeller = labeller(generation = 
    c("gen07" = "Generation 7",
      "gen31" = "Generation 31"))) 

connec_parallel_plot

# Export plot:

#ggsave("output_files/plots/Pleiotropy/corr_conn_parallelism.svg", connec_parallel_plot, units = "mm", width = 200, height = 100)

```
# Plot 2
```{r}
parallel_pleio_df %>% ggplot(aes(y=tau, x=factor(rep_count), fill=factor(rep_count))) +
  geom_boxplot() +
  #geom_violin() +
  scale_fill_manual(values =c("#fdd49e","#fc8d59","#ef6548","#d7301f","#7f0000")) +
  ylab("1-Tau") +
  xlab("Parallelism (No of replicates with expression changes)") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.ticks = element_line(color="black"),
    axis.text = element_text(size = 14, color = "black"),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position ="none",
    panel.border = element_rect(colour = "black",
                                fill=NA,
                                linewidth=0.5)) +
  facet_wrap(~generation) 

##

tau_parallel_plot <- parallel_pleio_df %>% ggplot(aes(y=tau, x=factor(rep_count), fill=factor(rep_count))) +
  #geom_boxplot() +
  geom_violin() +
  geom_boxplot(fill = "white", width=0.2, color="black", alpha=0.4, outlier.size = 0) +
  scale_fill_manual(values =c("#fdd49e","#fc8d59","#ef6548","#d7301f","#7f0000")) +
  ylab("1-Tau") +
  xlab("Parallelism (No of replicates with expression changes)") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.ticks = element_line(color="black"),
    axis.text = element_text(size = 14, color = "black"),
    strip.text.x.top = element_text(colour="black", face = "bold", size = 16),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    legend.position ="none",
    panel.border = element_rect(colour = "black",
                                fill=NA,
                                linewidth=0.5),
    strip.background = element_rect(colour="white",fill="white")) +
  facet_wrap(~ generation, labeller = labeller(generation = 
    c("gen07" = "Generation 7",
      "gen31" = "Generation 31"))) 

tau_parallel_plot

# Export plot:

#ggsave("output_files/plots/Pleiotropy/corr_tau_parallelism.svg", tau_parallel_plot, units = "mm", width = 200, height = 100)

```

# Wilkoxon tests - Connectivity 5 groups
```{r}
#### GENERATION 7:
# Filter F7:
connec_paralell_df_F7 <- parallel_pleio_df %>% dplyr::filter(generation == "gen07")
# Remove genes with no connectivity information:
connec_paralell_df_F7 <- connec_paralell_df_F7[!is.na(connec_paralell_df_F7$connectivity),]

wilcox_conn <- pairwise.wilcox.test(connec_paralell_df_F7$log10_connectivity, 
                                         connec_paralell_df_F7$rep_count, 
                                         p.adjust.method = "none")
# Results into a data-frame:
wilcox_test_df <- wilcox_conn$p.value
wilcox_test_df <- data.frame(expand.grid(dimnames(wilcox_test_df)), array(wilcox_test_df)) %>% na.omit()
colnames(wilcox_test_df) <- c("group1", "group2", "p_value")
wilcox_test_df$p_value <- as.numeric(wilcox_test_df$p_value)
# include generation:
wilcox_test_df$generation <- "Gen 7"
# Remove non-desired contrasts:
wilcox_test_df$contrast <- paste(wilcox_test_df$group1, wilcox_test_df$group2, sep=" VS ")
export_df_F7 <- wilcox_test_df %>% dplyr::filter(!contrast %in% c("plateau VS shared",
                                                "monotonic VS shared",
                                                "incomplete_reversed VS shared"))
####Generation 31

# Filter F7:
connec_paralell_df_F31 <- parallel_pleio_df %>% dplyr::filter(generation == "gen31")
# Remove genes with no connectivity information:
connec_paralell_df_F31 <- connec_paralell_df_F31[!is.na(connec_paralell_df_F31$connectivity),]

wilcox_conn <- pairwise.wilcox.test(connec_paralell_df_F31$log10_connectivity, 
                                         connec_paralell_df_F31$rep_count, 
                                         p.adjust.method = "none")
# Results into a data-frame:
wilcox_test_df <- wilcox_conn$p.value
wilcox_test_df <- data.frame(expand.grid(dimnames(wilcox_test_df)), array(wilcox_test_df)) %>% na.omit()
colnames(wilcox_test_df) <- c("group1", "group2", "p_value")
wilcox_test_df$p_value <- as.numeric(wilcox_test_df$p_value)
# include generation:
wilcox_test_df$generation <- "Gen 31"
# Remove non-desired contrasts:
wilcox_test_df$contrast <- paste(wilcox_test_df$group1, wilcox_test_df$group2, sep=" VS ")
export_df_F31 <- wilcox_test_df %>% dplyr::filter(!contrast %in% c("plateau VS shared",
                                                "monotonic VS shared",
                                                "incomplete_reversed VS shared"))

combined_conn_rep_wise_parallel_df <- rbind(export_df_F7, export_df_F31)


# Multiple test correction - FDR across 20 tests
combined_conn_rep_wise_parallel_df$padj_fdr <- p.adjust(combined_conn_rep_wise_parallel_df$p_value, method = "fdr", n = length(combined_conn_rep_wise_parallel_df$p_value))
# Include asterisks:
combined_conn_rep_wise_parallel_df <- combined_conn_rep_wise_parallel_df %>%
  mutate(significance = case_when(
    padj_fdr >= 0.05 ~ "n.s",          
    padj_fdr < 0.05 & padj_fdr > 0.01 ~ "*",  
    padj_fdr <= 0.01 & padj_fdr > 0.001 ~ "**", 
    padj_fdr <= 0.001 & padj_fdr > 0.0001 ~ "***" ,
    padj_fdr < 0.0001 ~ "****"))
# Export test results:
combined_conn_rep_wise_parallel_df$test <- "Wilcoxson rank sum test, two tailed"
# Include tested variable:
combined_conn_rep_wise_parallel_df$variable <- "log10(connectivity)"
combined_conn_rep_wise_parallel_df <- combined_conn_rep_wise_parallel_df %>% dplyr::select("variable","generation","contrast","p_value",
                         "padj_fdr","significance",
                         "test")
colnames(combined_conn_rep_wise_parallel_df) <- c("variable","generation","contrast" ,"p_value",
                         "padj_fdr","significancy_fdr","test")


```

# Wilkoxon tests - Tau 5 groups
```{r}
#### GENERATION 7:

# Filter F7:
tau_paralell_df_F7 <- parallel_pleio_df %>% dplyr::filter(generation == "gen07")
# Remove genes with no tau information:
tau_paralell_df_F7 <- tau_paralell_df_F7[!is.na(tau_paralell_df_F7$tau),]

wilcox_tau <- pairwise.wilcox.test(tau_paralell_df_F7$tau, 
                                         tau_paralell_df_F7$rep_count, 
                                         p.adjust.method = "none")
# Results into a data-frame:
wilcox_test_df <- wilcox_tau$p.value
wilcox_test_df <- data.frame(expand.grid(dimnames(wilcox_test_df)), array(wilcox_test_df)) %>% na.omit()
colnames(wilcox_test_df) <- c("group1", "group2", "p_value")
wilcox_test_df$p_value <- as.numeric(wilcox_test_df$p_value)
# include generation:
wilcox_test_df$generation <- "Gen 7"
# Remove non-desired contrasts:
wilcox_test_df$contrast <- paste(wilcox_test_df$group1, wilcox_test_df$group2, sep=" VS ")
export_df_F7 <- wilcox_test_df %>% dplyr::filter(!contrast %in% c("plateau VS shared",
                                                "monotonic VS shared",
                                                "incomplete_reversed VS shared"))

####Generation 31

# Filter F31:
tau_paralell_df_F31 <- parallel_pleio_df %>% dplyr::filter(generation == "gen31")
# Remove genes with no tau information:
tau_paralell_df_F31 <- tau_paralell_df_F31[!is.na(tau_paralell_df_F31$tau),]

wilcox_tau <- pairwise.wilcox.test(tau_paralell_df_F31$tau, 
                                         tau_paralell_df_F31$rep_count, 
                                         p.adjust.method = "none")
# Results into a data-frame:
wilcox_test_df <- wilcox_tau$p.value
wilcox_test_df <- data.frame(expand.grid(dimnames(wilcox_test_df)), array(wilcox_test_df)) %>% na.omit()
colnames(wilcox_test_df) <- c("group1", "group2", "p_value")
wilcox_test_df$p_value <- as.numeric(wilcox_test_df$p_value)
# include generation:
wilcox_test_df$generation <- "Gen 31"
# Remove non-desired contrasts:
wilcox_test_df$contrast <- paste(wilcox_test_df$group1, wilcox_test_df$group2, sep=" VS ")
export_df_F31 <- wilcox_test_df %>% dplyr::filter(!contrast %in% c("plateau VS shared",
                                                "monotonic VS shared",
                                                "incomplete_reversed VS shared"))


combined_tau_rep_wise_parallel_df <- rbind(export_df_F7, export_df_F31)

# Multiple test correction - FDR across 20 tests
combined_tau_rep_wise_parallel_df$padj_fdr <- p.adjust(combined_tau_rep_wise_parallel_df$p_value, method = "fdr", n = length(combined_tau_rep_wise_parallel_df$p_value))
# Include asterisks:
combined_tau_rep_wise_parallel_df <- combined_tau_rep_wise_parallel_df %>%
  mutate(significance = case_when(
    padj_fdr >= 0.05 ~ "n.s",          
    padj_fdr < 0.05 & padj_fdr > 0.01 ~ "*",  
    padj_fdr <= 0.01 & padj_fdr > 0.001 ~ "**", 
    padj_fdr <= 0.001 & padj_fdr > 0.0001 ~ "***" ,
    padj_fdr < 0.0001 ~ "****"))
combined_tau_rep_wise_parallel_df$padj_fdr <- format(combined_tau_rep_wise_parallel_df$padj_fdr, scientific = TRUE, digits = 2)
# Export test results:
combined_tau_rep_wise_parallel_df
combined_tau_rep_wise_parallel_df$test <- "Wilcoxson rank sum test, two tailed"
# Include tested variable:
combined_tau_rep_wise_parallel_df$variable <- "1-Tau"
combined_tau_rep_wise_parallel_df <- combined_tau_rep_wise_parallel_df %>% dplyr::select("variable","generation","contrast","p_value",
                         "padj_fdr","significance",
                         "test")
colnames(combined_tau_rep_wise_parallel_df) <- c("variable","generation","contrast" ,"p_value",
                         "padj_fdr","significancy_fdr","test")

```

# Export tests results:
```{r}
export_df <- rbind(combined_tau_rep_wise_parallel_df, combined_conn_rep_wise_parallel_df)

write_xlsx(export_df, "output_files/S7_wilcox_test_pleiotropy_and_parallellism.xlsx")

```

--- END ---